---
title: "Regression Example"
author: "Alan T. Arnholt"
date: "February 21, 2015"
output: html_document
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, fig.align = "center")
```

This example walks through creating a regression model using the `Boston` data frame from the `MASS` package.

```{r}
library(MASS)
names(Boston)
str(Boston)
```

We would like to create a model that predicts the median value of owner occupied homes (`medv`).  Start by viewing the relationship between `medv` and the other variables in the `Boston` data frame.  We will use the package `car`. If you do not have `car` installed on your machine, install it now.

```{r fig.height = 10, fig.width = 10}
library(car)  
scatterplotMatrix(~ medv + lstat + black + ptratio + tax + rad + dis + age + rm + nox + chas + indus + zn + crim, data = Boston)
```

Comment about scatter plot

## Simple Linear Model

```{r}
simple.lm <- lm(medv ~ lstat, data = Boston)
simple.lm
summary(simple.lm)
anova(simple.lm)
```

## Basic Regression Plots

```{r}
par(mfrow=c(2, 2))
plot(simple.lm)
par(mfrow=c(1, 1))
```

## Ggplot2 Graph

```{r}
library(ggplot2)
ggplot(data = Boston, aes(x = lstat, y = medv)) +
  geom_point() + 
  theme_bw() + 
  stat_smooth(method = "lm")
```

## Two Independent Variables

```{r}
twoiv.lm <- lm(medv ~ lstat + rm, data = Boston)
summary(twoiv.lm)
```

## Plane of Best Fit

```{r echo = FALSE}
library(scatterplot3d)
s3d <- scatterplot3d(x = Boston$lstat, y = Boston$rm, 
              z = Boston$medv, xlab = "Lower Status of Population", 
              ylab = "Average Number of Rooms", zlab = "Median Value of Owner Occupied Home",
              box = TRUE, pch = 20, color = "white",
              cex.symbols = 0.75, angle = 50, grid = FALSE)
s3d$plane3d(hsw.lm <- lm(medv ~ lstat + rm, data = Boston), 
            lty = "dotted", lty.box = "solid")
orig <- s3d$xyz.convert(x = Boston$lstat, y = Boston$rm, 
              z = Boston$medv)
plane <- s3d$xyz.convert(x = Boston$lstat, y = Boston$rm, 
              fitted(hsw.lm))
i.negpos <- 1 + (resid(hsw.lm) > 0)
segments(orig$x, orig$y, plane$x, plane$y,
         col = c("darkblue", "lightblue3")[i.negpos])
s3d$points3d(x = Boston$lstat, y = Boston$rm, 
              z = Boston$medv,
             col = c("darkblue", "lightblue3")[i.negpos],
             pch = 20)
```

## More Graphs Again

```{r}
par(mfrow=c(2, 2))
plot(twoiv.lm)
par(mfrow=c(1, 1))
```

## Forward and Backward Selection

The functions `add1` and `update` can be used to create a model with forward selection.

```{r ForwardSelection}
SCOPE <- (~ . + crim + zn + indus + chas + nox + rm + age + dis + rad + tax + ptratio + black + lstat)
mod.fs <- lm(medv ~ 1, data = Boston)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, . ~ . + lstat)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, . ~ . + rm)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, . ~ . + ptratio)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, . ~ . + dis)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, . ~ . + nox)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, . ~ . + chas)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, . ~ . + black)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, . ~ . + zn)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, . ~ . + crim)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, . ~ . + rad)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, . ~ . + tax)
add1(mod.fs, scope = SCOPE, test = "F")
summary(mod.fs)
```

## Backward Elimination

Backward elimination starts with all of the variables in the model and removes the least significant variable one at a time.

```{r modelall}
mod.be <- lm(medv ~ . , data = Boston)
drop1(mod.be, test = "F")  # single term deletions
mod.be <- update(mod.be, . ~ . - age)
drop1(mod.be, test = "F")  # single term deletions
mod.be <- update(mod.be, . ~ . - indus)
drop1(mod.be, test = "F")  # single term deletions
mod.be <- update(mod.be, . ~ . - age)
summary(mod.be)


## Criterion Based Selection

```{r}
library(leaps)
models <- regsubsets(medv ~ ., data = Boston, nvmax = 11)
summary(models)
R2adj <- summary(models)$adjr2
R2adj
which.max(R2adj)
MCP <- summary(models)$cp  # Mallow's C_p
MCP
BIC <- summary(models)$bic
BIC
which.min(BIC)
which.min(MCP)
```

Comment on model selected with $R^2_{adj}$, Mallow's $C_p$, and the Bayesian Information Criterion (BIC).

```{r fig.show='hold', fig.width = 10}
plot(models, scale = "adjr2")
plot(models, scale = "Cp")
plot(models, scale = "bic")
```

## Diagnostic Plots Again

```{r}
par(mfrow=c(2, 2))
plot(mod.be)
par(mfrow=c(1, 1))
```

## More Diagnostic Stuff

```{r}
library(car)
influenceIndexPlot(mod.be, id.n = 3)
outlierTest(mod.be)
influencePlot(mod.be)
InflMea <- influence.measures(mod.be)
summary(InflMea)  # Which observations are influential
```

## Transformations

```{r}
boxCox(mod.be)
boxCox(mod.be, lambda = seq(-0.1, 0.3, 0.01))
mod.log <- lm(log(medv) ~ crim + zn + chas + nox + rm + dis + rad + tax + ptratio + black + lstat, data = Boston)
boxCox(mod.log)
```

```{r}
par(mfrow=c(2, 2))
plot(mod.log)
par(mfrow=c(1, 1))
plot(mod.log, which = 1)
```

```{r}
library(car)
residualPlots(mod.log)
```

```{r}
mod2 <- lm(medv ~ crim + zn + chas + nox + rm + I(rm^2) + dis + rad + tax + ptratio + black + lstat + I(lstat^2), data = Boston)
summary(mod.log)
summary(mod2)
boxCox(mod2)
mod3 <- lm(medv ~ crim + zn + chas + nox + rm + I(rm^2) + dis + rad + tax + ptratio + black + lstat, data = Boston)
summary(mod3)
boxCox(mod3)
```