---
title: "Data Management"
author: "Alan T. Arnholt"
date: "January 2, 2015"
output: html_document
runtime: shiny
---

Start by loading your data set.  The url is broken up into two parts so that the 
R code will fit on the page.  The function `paste0()` is used to concatenate two
strings.

```{r comment = NA, cache=TRUE}
firstpart <- "http://www1.appstate.edu/~arnholta/classes/"
secondpart <- "PDS/DATAandCODEBOOKS/NESARC/NESARC_pds.RData"
site <-paste0(firstpart, secondpart) 
con <- url(site)
load(con)
NESARC[1:6, 1:10] # show first six rows and first 10 columns
```

### Using the `dplyr` package
The package `dplyr` must first be installed before it can be used.  There are two ways one can install the package.  The easiest way to install `dplyr`, is to type 
```{r eval=FALSE}
install.packages("dplyr")
```
at the `R` prompt.  If you want the lastest version of `dplyr`, you can install the development version from GitHub provided you have the required tools to build `R` packages installed on your machine (`Rtools` for Windows users, `XCode` for Mac users, etc.).  

```{r eval=FALSE}
if (packageVersion("devtools") < 1.6) {
  install.packages("devtools")
}
devtools::install_github("hadley/lazyeval")
devtools::install_github("hadley/dplyr")
```

### Renaming Variables
Renaming the variables `IDNUM`, `ETOTLCA2`, and `ETHRACE2A` to `UniqueID`, `EthanolConsumption`, and `Ethnicity` is done with the `rename` function.  Although it is not a requirement, the data frame `NESARC` is converted to a `data frame tbl` per the advice given in the `dplyr` vignette.  The `rename` function takes a data frame or data frame tbl as its first argument while the syntax `new_name = old_name` is used as the second argument.

```{r comment = NA, message = FALSE}
library(dplyr)
# Create a data frame tbl see ?tbl_df
NESARCtbl <- tbl_df(NESARC)
NESARC <- rename(NESARCtbl, UniqueID = IDNUM, EthanolConsumption = ETOTLCA2, Ethnicity = ETHRACE2A)
NESARC[1:6, 1:5]
```

### Coding missing values
Note that the variable `S3AQ3B1` uses a 9 to record `unknown` for smoking frequency.

```{r comment = NA}
NESARC$S3AQ3B1[NESARC$S3AQ3B1 == 9] <- NA
summary(NESARC$S3AQ3B1)  # Note that 9 still appears
NESARC$S3AQ3B1 <- factor(NESARC$S3AQ3B1)[, drop = TRUE]
summary(NESARC$S3AQ3B1)  # Unused level no longer appears
```

I hate to look back at the code book to figure out what the heck the numbers represent in this case!  One of the great things about the `R` language is how it handles a wide variety of data types.  In this case, the variable `S3AQ3B1` is a factor with numeric labels for the levels of the factor.  To change the numeric labels to text type the following:

```{r comment = NA}
NESARC$S3AQ3B1 <- factor(NESARC$S3AQ3B1, labels = c("Every Day", "5 to 6 Days/week", "3 to 4 Days/week", "1 to 2 Days/week", "2 to 3 Days/month", "Once a month or less"))
summary(NESARC$S3AQ3B1)
xtabs(~S3AQ3B1, data = NESARC) # Note how the NA's are not printed
```

The following `ggplot2` code is used to graph the values in `S3AQ3B1`.

```{r message = FALSE, fig.width = 10, fig.wdith = 5}
library(ggplot2)
ggplot(data = NESARC, aes(x = S3AQ3B1)) + 
  geom_bar(fill = "lightgray") + 
  labs(x = "Smoking Frequency") +
  theme_bw() 
```

If you do not wnat to see the `NA`'s use
```{r message = FALSE, fig.width = 10, fig.wdith = 5}
ggplot(data = na.omit(NESARC[ , "S3AQ3B1", drop = FALSE]), aes(x = S3AQ3B1)) + 
  geom_bar(fill = "lightgray") + 
  labs(x = "Smoking Frequency") +
  theme_bw() 
```

The variable `S4AQ4A1` represents unknown values with a 9, `No` values with a 2, and `Yes` values with a 1.  

```{r comment = NA}
summary(NESARC$S4AQ4A1)
```

To recode the NA's as `No` enter
```{r comment = NA}
NESARC$S4AQ4A1[is.na(NESARC$S4AQ4A1)] <- 2
summary(NESARC$S4AQ4A1)
NESARC$S4AQ4A1 <- factor(NESARC$S4AQ4A1)
summary(NESARC$S4AQ4A1)
```

### Collapsing Categories
Consider the varaible `S1Q6A` with 14 levels recording the highest level of education of the participant.  To collapse the categories in to dichotomous variable indicating the presence of a high school degree use the `ifelse` function. The levels `1`, `2`, `3`, `4`, `5`, `6`, and `7` of the variable `S1Q6A` correspond to education levels less than completing high school.

```{r comment = NA}
NESARC$HS_DEGREE <- factor(ifelse(NESARC$S1Q6A %in% c("1", "2", "3", "4", "5", "6", "7"), "No", "Yes"))
summary(NESARC$HS_DEGREE)
```

### Creating a Factor from a Numeric Vector
To create a factor with three levels (young adults, adults, older adults) using the variable `AGE` (a numeric vector) use the function `cut`.

```{r comment = NA}
NESARC$AGEfac <- cut(NESARC$AGE, breaks = c(18, 30, 50, Inf), labels = c("Young Adult", "Adult", "Older Adult"), include.lowest = TRUE)
summary(NESARC$AGEfac)
```

Consider collapsing the numeric vector `S3AQ3C1` (usual quantity when cigarettes smoked) into a categorical variable with 5 levels.

```{r comment = NA}
NESARC$S3AQ3C1fac <- cut(NESARC$S3AQ3C1, breaks = c(0, 5, 10, 15, 20, 100), include.lowest = TRUE)
summary(NESARC$S3AQ3C1fac)
```

### Aggregating Variables using `ifelse`
Consider creating a new variable `DepressLife` which is `Yes` if the variable `MAJORLIFE` is a 1 or if `DYSLIFE` is a 1.

```{r comment = NA}
NESARC$DepressLife <- factor(ifelse( (NESARC$MAJORDEPLIFE == 1 | NESARC$DYSLIFE == 1), "Yes", "No"))
summary(NESARC$DepressLife)
```

As another example, consider creating a variable `PPpanic` for past and present panic disorders.  The variables `APANDX12` and `APANDXP12` record a panic disorder within the last 12 months and prior to the last 12 months with agoraphobia, respectively.  The variables `PANDX12` and `PANDXP12` record a panic disorder within the last 12 months and prior to the last 12 months without agoraphobia, respectively.

```{r comment = NA}
NESARC$PPpanic <- factor(ifelse( (NESARC$APANDX12 == 1 | NESARC$APANDXP12 == 1 | NESARC$PANDX12 == 1 | NESARC$PANDXP12 == 1 ), "Yes", "No"))
summary(NESARC$PPpanic)
```

Consider a new variable `AllDeprSymp` that records whether a subject exhibits all depression symptoms or not.  There are 19 depression symptoms recorded in the variables `S4AQ4A1` - `S4AQ4A19`.  Note that the logical operator `&` is used to test if all depression symptoms are present.

```{r comment = NA}
NESARC$AllDeprSymp <- factor(ifelse( (NESARC$S4AQ4A1 == 1 & NESARC$S4AQ4A2 == 1 & NESARC$S4AQ4A3 == 1 & NESARC$S4AQ4A4 == 1 & NESARC$S4AQ4A5 == 1 & NESARC$S4AQ4A6 == 1 & NESARC$S4AQ4A7 == 1 & NESARC$S4AQ4A8 == 1 & NESARC$S4AQ4A9 == 1 & NESARC$S4AQ4A10 == 1 & NESARC$S4AQ4A11 == 1 & NESARC$S4AQ4A12 == 1 & NESARC$S4AQ4A13 == 1 & NESARC$S4AQ4A14 == 1 & NESARC$S4AQ4A15 == 1 & NESARC$S4AQ4A16 == 1 & NESARC$S4AQ4A17 == 1 & NESARC$S4AQ4A18 == 1 & NESARC$S4AQ4A19 == 1), "Yes", "No"))
summary(NESARC$AllDeprSymp)
```

### Creating a Composite Factor

To create a variable `NumDepSym` that records the number of depression symptoms among the variables `S4AQ4A1` - `S4AQ4A19`, create numeric variables from the factors `S4AQ4A1` - `S4AQ4A19` and store the results in `S4AQ4A1num` - `S4AQ4A19num`.  Next, recode the levels of `No` (2) and `Unknown` (3) to 0s for varaibles `S4AQ4A1num` - `S4AQ4A19num`.  The code is rather lengthy and I welcome other approaches.  The template given in the PDS IBook link will not work with this data because the variables are factors not numeric vectors.  You can not add factors!

```{r comment = NA}
summary(NESARC$S4AQ4A1)
levels(NESARC$S4AQ4A1)
table(as.numeric(NESARC$S4AQ4A1))
NESARC$S4AQ4A1num <- as.numeric(NESARC$S4AQ4A1)
xtabs(~S4AQ4A1num, data = NESARC)
NESARC$S4AQ4A1num[NESARC$S4AQ4A1num == 2 | NESARC$S4AQ4A1num == 3] <- 0
xtabs(~S4AQ4A1num, data = NESARC)
NESARC$S4AQ4A2num <- as.numeric(NESARC$S4AQ4A2)
NESARC$S4AQ4A2num[NESARC$S4AQ4A2num == 2 | NESARC$S4AQ4A2num == 3] <- 0
NESARC$S4AQ4A3num <- as.numeric(NESARC$S4AQ4A3)
NESARC$S4AQ4A3num[NESARC$S4AQ4A3num == 2 | NESARC$S4AQ4A3num == 3] <- 0
NESARC$S4AQ4A4num <- as.numeric(NESARC$S4AQ4A4)
NESARC$S4AQ4A4num[NESARC$S4AQ4A4num == 2 | NESARC$S4AQ4A4num == 3] <- 0
NESARC$S4AQ4A5num <- as.numeric(NESARC$S4AQ4A5)
NESARC$S4AQ4A5num[NESARC$S4AQ4A5num == 2 | NESARC$S4AQ4A5num == 3] <- 0
NESARC$S4AQ4A6num <- as.numeric(NESARC$S4AQ4A6)
NESARC$S4AQ4A6num[NESARC$S4AQ4A6num == 2 | NESARC$S4AQ4A6num == 3] <- 0
NESARC$S4AQ4A7num <- as.numeric(NESARC$S4AQ4A7)
NESARC$S4AQ4A7num[NESARC$S4AQ4A7num == 2 | NESARC$S4AQ4A7num == 3] <- 0
NESARC$S4AQ4A8num <- as.numeric(NESARC$S4AQ4A8)
NESARC$S4AQ4A8num[NESARC$S4AQ4A8num == 2 | NESARC$S4AQ4A8num == 3] <- 0
NESARC$S4AQ4A9num <- as.numeric(NESARC$S4AQ4A9)
NESARC$S4AQ4A9num[NESARC$S4AQ4A9num == 2 | NESARC$S4AQ4A9num == 3] <- 0
NESARC$S4AQ4A10num <- as.numeric(NESARC$S4AQ4A10)
NESARC$S4AQ4A10num[NESARC$S4AQ4A10num == 2 | NESARC$S4AQ4A10num == 3] <- 0
NESARC$S4AQ4A11num <- as.numeric(NESARC$S4AQ4A11)
NESARC$S4AQ4A11num[NESARC$S4AQ4A11num == 2 | NESARC$S4AQ4A11num == 3] <- 0
NESARC$S4AQ4A12num <- as.numeric(NESARC$S4AQ4A12)
NESARC$S4AQ4A12num[NESARC$S4AQ4A12num == 2 | NESARC$S4AQ4A12num == 3] <- 0
NESARC$S4AQ4A13num <- as.numeric(NESARC$S4AQ4A13)
NESARC$S4AQ4A13num[NESARC$S4AQ4A13num == 2 | NESARC$S4AQ4A13num == 3] <- 0
NESARC$S4AQ4A14num <- as.numeric(NESARC$S4AQ4A14)
NESARC$S4AQ4A14num[NESARC$S4AQ4A14num == 2 | NESARC$S4AQ4A14num == 3] <- 0
NESARC$S4AQ4A15num <- as.numeric(NESARC$S4AQ4A15)
NESARC$S4AQ4A15num[NESARC$S4AQ4A15num == 2 | NESARC$S4AQ4A15num == 3] <- 0
NESARC$S4AQ4A16num <- as.numeric(NESARC$S4AQ4A16)
NESARC$S4AQ4A16num[NESARC$S4AQ4A16num == 2 | NESARC$S4AQ4A16num == 3] <- 0
NESARC$S4AQ4A17num <- as.numeric(NESARC$S4AQ4A17)
NESARC$S4AQ4A17num[NESARC$S4AQ4A17num == 2 | NESARC$S4AQ4A17num == 3] <- 0
NESARC$S4AQ4A18num <- as.numeric(NESARC$S4AQ4A18)
NESARC$S4AQ4A18num[NESARC$S4AQ4A18num == 2 | NESARC$S4AQ4A18num == 3] <- 0
NESARC$S4AQ4A19num <- as.numeric(NESARC$S4AQ4A19)
NESARC$S4AQ4A19num[NESARC$S4AQ4A19num == 2 | NESARC$S4AQ4A19num == 3] <- 0
NESARC$NumDepSym <- NESARC$S4AQ4A1num + NESARC$S4AQ4A2num + NESARC$S4AQ4A3num + NESARC$S4AQ4A4num + NESARC$S4AQ4A5num + NESARC$S4AQ4A6num + NESARC$S4AQ4A7num + NESARC$S4AQ4A8num + NESARC$S4AQ4A9num + NESARC$S4AQ4A10num + NESARC$S4AQ4A11num + NESARC$S4AQ4A12num + NESARC$S4AQ4A13num + NESARC$S4AQ4A14num + NESARC$S4AQ4A15num + NESARC$S4AQ4A16num + NESARC$S4AQ4A17num + NESARC$S4AQ4A18num + NESARC$S4AQ4A19num
xtabs(~NumDepSym, data = NESARC)
```

### Subsetting Your Data
Three different approaches to subsetting data will be given.  The first approach is to use the `dplyr` function `filter`; the second approach is to use indices; and the third approach is to use the function `subset`.  Consider creating a subset of the `NESARC` data set where a person indicates

1. He/she has smoked over 100 cigarettes (`S3AQ1A == 1`)
2. He/she has smoked in the past year (`CHECK321 == 1`)
3. He/she has typically smoked every day over the last year (`S3AQ3B1 == 1`)
4. He/she is less than or equal to 25 years old (`AGE <= 25`)


The first approach uses the `filter` function with the `%>%` function.
```{r comment = NA}
NESARCsub1 <- NESARC %>%
  filter(S3AQ1A == 1 & CHECK321 == 1 & S3AQ3B1 == "Every Day" & AGE <= 25)
dim(NESARCsub1)
```

The second approach uses standard indexing.
```{r comment = NA}
NESARCsub2 <- NESARC[NESARC$S3AQ1A == 1 & NESARC$CHECK321 == 1 & NESARC$S3AQ3B1 == "Every Day" & NESARC$AGE <= 25  & !is.na(NESARC$S3AQ1A == 1 & NESARC$CHECK321 == 1 & NESARC$S3AQ3B1 == "Every Day" & NESARC$AGE <= 25), ]
dim(NESARCsub2)
```

The third approach uses the `subset` function.
```{r comment = NA}
NESARCsub3 <- subset(NESARC, subset = S3AQ1A == 1 & CHECK321 == 1 & S3AQ3B1 == "Every Day" & AGE <= 25)
dim(NESARCsub3)
```

### ANOVA example

More examples to appear soon...

```{r comment = NA,fig.align='center'}
my.aov <- aov(S3AQ3C1 ~ TAB12MDX, data = subset(NESARC, S3AQ3C1 < 99)) # Note S3AQ3C1 == 99 is NA
summary(my.aov)     
TukeyHSD(my.aov)
plot(TukeyHSD(my.aov))
```

### Graphing a Numeric Variable with `ggplot2`

```{r comment = NA,fig.align='center'}
ggplot(data = NESARC, aes(x = AGE)) + 
  geom_histogram(binwidth = 5, fill = "lightgray") + 
  theme_bw()  
ggplot(data = NESARC, aes(x = AGE, y = ..density..)) + 
  geom_histogram(binwidth = 5, fill = "lightgray") + 
  theme_bw()  
ggplot(data = NESARC, aes(x = AGE, y = ..density..)) + 
  geom_histogram(binwidth = 5, fill = "lightgray") + 
  theme_bw() + 
  geom_density(fill = "yellow", alpha = 0.2)
```



```{r label = "Mini", comment = NA, fig.align = 'center'}
MINI <- NESARC %>%
  select(S1Q24FT, S1Q24IN, S1Q24LB, SEX) %>%
  filter(S1Q24FT < 99, S1Q24IN < 99, S1Q24LB < 999) %>%
  mutate(Inches = (S1Q24FT*12 + S1Q24IN), Sex = factor(SEX, labels =c("Male", "Female")), Weight = S1Q24LB) %>%
  select(Sex, Inches, Weight)
MINI
```

```{r label = "WtHist", comment = NA, fig.align = "center"}
ggplot(data = MINI, aes(x = Inches)) + 
  geom_histogram(binwidth = 1, fill = "peachpuff") + 
  theme_bw()
ggplot(data = MINI, aes(x = Inches)) + 
  geom_density(aes(fill = Sex, color = Sex)) + 
  theme_bw() + 
  facet_grid(Sex ~ .) 
tapply(MINI$Inches, MINI$Sex, median)  
```

### Some Bivariate Graphs with `ggplot2`

```{r label = "LastTwo", comment = NA, fig.align = 'center'}
ggplot(data = MINI, aes(x = Inches, y = Weight, color = Sex)) + 
  geom_point(alpha = 0.1) + 
  stat_smooth(method = "lm") + 
  theme_bw() + 
  labs(y = "Weight (pounds)")
ggplot(data = MINI, aes(x = Inches, y = Weight)) + 
  geom_point(alpha = 0.1, aes(color = Sex)) +
  stat_smooth(method = "lm", aes(color = Sex)) + 
  facet_grid(Sex ~ .) + 
  theme_bw() + 
  labs(y = "Weight (pounds)")
```

### Graphing with `ggvis`

Move the mouse over a point in the next graph to get information on that point.  This will only work when there is an `R` process running.

```{r comment = NA, message = FALSE, warning = FALSE, fig.align='center'}
all_values <- function(x) {
  if(is.null(x)) return(NULL)
  paste0(names(x), ": ", format(x), collapse = "<br />")
}
library(ggvis)
MINI %>%
  ggvis(~Inches, ~Weight, fill = ~Sex) %>%
  layer_points() %>%
  group_by(Sex) %>%
  layer_model_predictions(formula = Weight ~ Inches, model = "lm", se = TRUE) %>%
  add_tooltip(all_values, "hover")
```

```{r}
MINI %>%
 ggvis(~Weight, fill = ~Sex) %>%
 group_by(Sex) %>%
 layer_densities(stroke = ~Sex, fill = ~Sex,   adjust = input_slider(0.1, 5, label = "Hold Down the Mouse and Move Me"))
```