---
title: "Data Management Introduction"
author: "Alan T. Arnholt"
date: "February 9, 2015"
output: html_document
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, fig.align = "center")
library(dplyr)
library(ggplot2)
library(PASWR2)
```

```{r}
str(EPIDURAL)
epi <- EPIDURAL %>%
  rename(Doctor = doctor, Weight = kg, Height = cm, Ease = ease, Treatment = treatment, OC = oc) %>%
  mutate(BMI = Weight/(Height/100)^2) %>%
  select(Doctor, Ease, Treatment, OC, BMI)
head(epi)
```

## Return rows with matching conditions (filter)

```{r}
epiAB <- epi %>%
  filter(BMI <= 30 & (Doctor == "Dr. A" | Doctor == "Dr. B"))
epiAB
epiCD <- epi %>%
  filter(BMI <= 30 & !(Doctor == "Dr. A" | Doctor == "Dr. B"))
epiCD
```

How many patients have a `BMI` between 20 and 25 inclusive and were treated by `Dr. A`?

```{r}
epiN <- epi %>%
  filter(BMI <= 25 & BMI >= 20 & Doctor == "Dr. A")
epiN
DN <- dim(epiN)
DN
```

There are `r DN[1]` patients that have a `BMI` between 20 and 25 inclusive and 
were treated by `Dr. A`.

## Problems

Use the data frame `quine` from the `MASS` package to answer all questions.

1.  How many aboriginal average learning females in grade "F0" have missed more than
5 days of school? (3)

```{r}
library(dplyr)
library(MASS)
qu <- quine %>%
  filter(Eth == "A" & Sex == "F" & Age == "F0" & Lrn == "AL" & Days > 5)
qu
n <- dim(qu)[1]
n
```

There are `r n`  aboriginal average learning females in grade "F0" have missed more than
5 days of school.

2.  How many non-aboriginal males in grade "F0" have never missed any school? (2)

```{r}
qu <- quine %>%
  filter(Eth == "N" & Sex == "M" & Age == "F0" & Days == 0)
qu
n <- dim(qu)[1]
n
```

There are `r n` non-aboriginal males in grade "F0" that have never missed any school.

3.  What is the average number of missed school days for non-aboriginal average learning males? (13.36364)

```{r}
qu <- quine %>%
  filter(Eth == "N" & Lrn == "AL" & Sex == "M")
qu
MeanDays <- mean(qu$Days)
MeanDays
TT <- tapply(quine$Days, list(quine$Eth, quine$Lrn, quine$Sex), mean)
TT
TT[, , "M"][2, 1]
```

The average number of missed school days for non-aboriginal average learning males is `r MeanDays`.

Use the following code to create the data frame `DF`.

```{r}
set.seed(37)
fac1 <- factor(sample(letters[1:4], 100, replace = TRUE))
fac2 <- factor(sample(LETTERS[1:2], 100, replace = TRUE))
sex <- factor(sample(c("Male", "Female"), 100, replace = TRUE))
DF <- data.frame(fac1, fac2, sex)
rm(fac1, fac2, sex)
head(DF)
```

4.  What percent of level `B` (`fac2`) are `Female`? (53.65854%)

```{r}
T1 <- xtabs(~ fac2 + sex, data = DF)
T1
prop.table(T1, 1)
# or in one line
ans <- round(prop.table(xtabs(~ fac2 + sex, data = DF), 1)[2, 1]*100, 2)
ans
```

`r round(ans, 2)`% of level `B` (`fac2`) are `Female`.
