---
title: "Class Outline"
author: "Alan T. Arnholt"
date: "January 20, 2015"
output: html_document
---

1. Zotero --- Introduce --- If Git is working start using Zotero and prepare notebook based on directions in PDS (pg. 13) for Thursday.

2. If you want to listen to videos during class please bring headphones.

3. Friday I learned a little about the MAC setup...

```{r engine = 'bash', comment = NA}
git version
```

First check that students have told Git who they are.  Run the following command at the shell:

```{r engine = 'bash', comment = NA}
git config --list
```

Must have at a minimum a `user.name` and `user.email`.  If this is not set up, run the following and make sure the email the student used to register  with GitHub is the email specified in the `user.email`.

```{r engine = 'bash', comment = NA, eval = FALSE}
git config --global user.name "Student Name"
git config --global user.email "student@appstate.edu"
```

When the `user.name` and `user.email` are entered run

```{r engine = 'bash', comment = NA, eval = FALSE}
git config --list
```

## Turn on the credential helper

**Do not do this on the lab machines!** This is only for your personal machines.

### Windows
In the shell, type
```{r engine = 'bash', eval = FALSE}
git config --global credential.helper wincred
```

### MAC
In the terminal, type
```{r engine ="bash", comment = NA, eval = FALSE}
git config --global credential.helper osxkeychain
```

## Trigger a username / password challenge

Fork the repo https://github.com/alanarnholt/STT3530 to your public repository (@username).  Change or add something to the README file.  Once the file has been changed, enter the following in the shell:

```{r engine = 'bash', eval = FALSE}
git add .
git commit -m"Some message"
git push
```

You will be asked one last time for your username and password.  

## See if RStudio works now

Make another change (say to the README) stage and commit the changes using RStudio.  Make sure to write a commit message then push your changes to your public repository.  

## Last 15 Minutes of Class

* Fork the repository https://github.com/STT2810-SP15/STT2810ClassRepoSP15 to your public repository.
* Clone the repository to your local machine using RStudio by following these instructions:
1. Copy the clone URL to the clipboard.
2. Click `File > New Project > Version Control > Git`
3. Paste the clone URL (`https://github.com/YourUSERNAME/STT2810ClassRepoSP15.git`) in the `Repository URL:` box.
4. Type a name (suggestion `UserNameSTT2810`) in the `Project directory name:` box.
5. Change if needed the location in the `Create project as subdirectory of:` box.
6. Click the `Create Project` box.  You should now have a local copy of the forked repository on your local machine.  Congratulations!  

* Set the upstream remote in your fork to this repository with the command
```{r engine = "bash", eval = FALSE, comment = NA}
git remote add upstream https://github.com/STT2810-SP15/STT2810ClassRepoSP15.git
```

* Verify with
```{r engine = 'bash', eval = FALSE}
git remote -v      
origin  https://github.com/alanarnholt/STT2810ClassRepoSP15.git (fetch)    
origin	https://github.com/alanarnholt/STT2810ClassRepoSP15.git (push)    
upstream	https://github.com/STT2810-SP15/STT2810ClassRepoSP15.git (fetch)     
upstream	https://github.com/STT2810-SP15/STT2810ClassRepoSP15.git (push)

```
* To obtain updates from the upstream repository type
```
git pull upstream master
```
* If there are conflicts you will need to resolve them before proceeding.