---
title: "Some ggvis Graphs"
author: "Arnholt, Alan"
date: "January 4, 2015"
output: html_document
runtime: shiny
---

Start by loading your data set.  The url is broken up into two parts so that the 
R code will fit on the page.  The function `paste0()` is used to concatenate two
strings.

```{r comment = NA, cache=TRUE}
firstpart <- "http://www1.appstate.edu/~arnholta/classes/"
secondpart <- "PDS/DATAandCODEBOOKS/NESARC/NESARC_pds.RData"
site <-paste0(firstpart, secondpart) 
con <- url(site)
load(con)
NESARC[1:6, 1:10] # show first six rows and first 10 columns
```

### Using the `dplyr` package
The package `dplyr` must first be installed before it can be used.  There are two ways one can install the package.  The easiest way to install `dplyr`, is to type 
```{r eval=FALSE}
install.packages("dplyr")
```
at the `R` prompt.  If you want the lastest version of `dplyr`, you can install the development version from GitHub provided you have the required tools to build `R` packages installed on your machine (`Rtools` for Windows users, `XCode` for Mac users, etc.).  

```{r eval=FALSE}
if (packageVersion("devtools") < 1.6) {
  install.packages("devtools")
}
devtools::install_github("hadley/lazyeval")
devtools::install_github("hadley/dplyr")
```

```{r comment = NA, message = FALSE}
library(dplyr)
# Create a data frame tbl see ?tbl_df
NESARC <- tbl_df(NESARC)
NESARC[1:6, 1:5]
```

```{r label = "Mini", comment = NA, fig.align = 'center'}
MINI <- NESARC %>%
  select(S1Q24FT, S1Q24IN, S1Q24LB, SEX) %>%
  filter(S1Q24FT < 99, S1Q24IN < 99, S1Q24LB < 999) %>%
  mutate(Inches = (S1Q24FT*12 + S1Q24IN), Sex = factor(SEX, labels =c("Male", "Female")), Weight = S1Q24LB) %>%
  select(Sex, Inches, Weight)
MINI
```


### Graphing with `ggvis`

Move the mouse over a point in the next graph to get information on that point.  This will only work when there is an `R` process running.

```{r comment = NA, message = FALSE, warning = FALSE, fig.align='center'}
all_values <- function(x) {
  if(is.null(x)) return(NULL)
  paste0(names(x), ": ", format(x), collapse = "<br />")
}
library(ggvis)
MINI %>%
  ggvis(~Inches, ~Weight, fill = ~Sex) %>%
  layer_points() %>%
  group_by(Sex) %>%
  layer_model_predictions(formula = Weight ~ Inches, model = "lm", se = TRUE) %>%
  add_tooltip(all_values, "hover")
```



```{r}
MINI %>%
 ggvis(~Weight, fill = ~Sex) %>%
 group_by(Sex) %>%
 layer_densities(stroke = ~Sex, fill = ~Sex,   adjust = input_slider(0.1, 5, label = "Hold Down the Mouse and Move Me"))
```


