---
title: "Data Management Stuff"
author: ""
date: "Spring 2015"
output: html_document
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, fig.align = "center")
library(dplyr)
library(ggplot2)
library(descr)
```

We will consider variables `H1SU1-H1SU3` from the `AddHealth` data set.  We will recode
responses of 6, 8, and 9 for `H1SU1` as `NA`. (Discuss whether this is appropriate)

```{r}
library(PDS)
xtabs(~H1SU1, data = AddHealth)
AddHealth$H1SU1[AddHealth$H1SU1==6 | AddHealth$H1SU1==8 | AddHealth$H1SU1==9] <- NA
xtabs(~H1SU1, data = AddHealth)
AddHealth$H1SU1 <- factor(AddHealth$H1SU1, labels = c("No", "Yes"))[, drop = TRUE]
xtabs(~H1SU1, data = AddHealth)
summary(AddHealth$H1SU1)
freq(AddHealth$H1SU1)
freq(AddHealth$H1SU1, main = "Did you ever seriously think about committing suicide?", 
     col = c("skyblue", "gray"))
ggplot(data = AddHealth, aes(x = H1SU1, fill = H1SU1)) + 
  geom_bar() + 
  labs(title = "Did you ever seriously think about committing suicide?", x = "") +
  theme_bw() 
```


Question `H1SU2` asks the 821 respondents from `H1SU1` "how many times did you attempt suicide?"  Response of
a 7 should be recoded to `NA`.

```{r}
xtabs(~H1SU2, data = AddHealth)
AddHealth$H1SU2[AddHealth$H1SU2==7] <- NA
xtabs(~H1SU2, data = AddHealth)
AddHealth$H1SU2 <- factor(AddHealth$H1SU2)[, drop = TRUE]
xtabs(~H1SU2, data = AddHealth)
AddHealth$H1SU2 <- factor(AddHealth$H1SU2, 
                          labels = c("0 times", "1 time", "2 or 3 times", 
                                     "4 or 5 times", ">= 6 times", "refused", 
                                     "don't know"))
summary(AddHealth$H1SU2)
xtabs(~H1SU2, data = AddHealth)
NAS <- sum(xtabs(~H1SU2, data = AddHealth)[2:5])  # number that attempted suicide
NAS
freq(AddHealth$H1SU2, main = "How many times did you actually attempt suicide?")
ggplot(data = AddHealth, aes(x = H1SU2)) + 
  geom_bar(fill = "maroon") + 
  labs(title = "How many times did you actually attempt suicide?", x= "") +
  theme_bw() 
ggplot(data = na.omit(AddHealth[ , "H1SU2", drop = FALSE]), aes(x = H1SU2)) + 
  geom_bar(fill = "plum") + 
  labs(title = "How many times did you actually attempt suicide?", x = "") +
  theme_bw() 
ggplot(data = na.omit(AddHealth[ , "H1SU2", drop = FALSE]), aes(x = H1SU2, fill = H1SU2)) + 
  geom_bar() + 
  labs(title = "How many times did you actually attempt suicide?", x = "") +
  theme_bw() 
```

Question `H1SU3` asks of the `r NAS` that attempted suicide if they had to be treated by either a nurse or doctor.  The response of 7 should be recoded to `NA`.

```{r}
AddHealth$H1SU3[AddHealth$H1SU3 == 7] <- NA
AddHealth$H1SU3 <- factor(AddHealth$H1SU3, labels = c("No", "Yes"))[, drop = TRUE]
xtabs(~H1SU3, data = AddHealth)
freq(AddHealth$H1SU3, main = "Did you require treatment by either a nurse of doctor?")
ggplot(data = na.omit(AddHealth[ , "H1SU3", drop = FALSE]), aes(x = H1SU3, fill = H1SU3)) + 
  geom_bar() + 
  labs(title = "Did you require treatment by either a nurse of doctor?", x = "") +
  theme_bw() 
```

## Numeric Variables

Imagine one wants to study GDP and military personnel.  The gapminder code-book indicates the variable
`armedforcesrate` measures the military personnel as a percentage of the total labor force.  However, no date is given for the data collection.  Using the `WDI` package we decide to run the following:

```{r}
library(WDI)
WDIsearch("armed forces")
WDIsearch("GDP")
AFP <- WDI(country = "all", indicator = c("MS.MIL.TOTL.TF.ZS", "NY.GDP.PCAP.PP.KD"),
    start = 2005, end = 2005, extra = FALSE, cache = NULL)
head(AFP)
GAP <- inner_join(x = gapminder, y = AFP)
GAP <- GAP %>%
  rename(AFR2005 = MS.MIL.TOTL.TF.ZS, GDP2005 = NY.GDP.PCAP.PP.KD) %>%
  select(armedforcesrate, AFR2005, country, GDP2005)
head(GAP)
ggplot(data = GAP, aes(x = AFR2005)) + 
  geom_density(fill = "pink") + 
  theme_bw()
ggplot(data = GAP, aes(x = GDP2005)) + 
  geom_density(fill = "pink") + 
  theme_bw()
ggplot(data = GAP, aes(log(AFR2005), log(GDP2005))) + 
  geom_point(color = "red") + 
  theme_bw() +
  labs(x = "log-Armed Forces Rate (2005)", y = "log-Gross Domestic Product (2005) per Capita")
cor(log(GAP$AFR2005), log(GAP$GDP2005), use = "complete.obs")
```
