---
title: "Data Management Template"
author: "Alan T. Arnholt"
date: "February 16, 2015"
output: html_document
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, fig.align = "center")
```

### Purpose of Study: (Note: This is the last paragraph of the research proposal used to remind the reader of the research objectives.)

The present study will examine young adults from the National Epidemiologic Survey of Alcohol and Related Conditions (NESARC). The goals of the analysis will include 1) establishing the relationship between major depression and nicotine dependence; and 2) determining whether or not the relationship between major depression and nicotine dependence exists above and beyond smoking quantity. 

### Variables:

Variables from NESARC that will be used include: `MAJORDEPLIFE` (has subject experienced a major depression?), `CHECK321` (has subject smoked cigarettes in past 12 months?), `AGE` (age of subject), `TAB12MDX` (tobacco dependence past 12 months), `S3AQ3B1` (usual frequency when cigarettes smoked), `ETHRACE2A` (ethnicity of subject), `SEX` (gender of subject), and `S3AQ3C1` (usual smoking quantity of cigarettes when smoked).  ￼

### Data Management

First, the data is placed on the search path using the `PDS` package.
Next, a subset of people 25 or younger who smoked in the last 12 months is created using the `filter` function from `dplyr`. Note that `CHECK321 == 1` is used to see if subject has smoked in the past 12 months and `!is.na()` is used to make sure the subset does not contain `NA`s for the chosen variables.  Last, the variables of interest are selected and stored in the data frame `nesarc.sub` using the `select` function from the `dplyr` package.

```{r}
library(PDS)
library(dplyr)
nesarc.sub <- NESARC %>% 
  filter(!is.na(CHECK321) & !is.na(AGE) & CHECK321 ==1 & AGE <= 25)
dim(nesarc.sub)
nesarc.sub <- nesarc.sub %>% 
  select(MAJORDEPLIFE, CHECK321, AGE, TAB12MDX, S3AQ3B1, S3AQ3C1, ETHRACE2A, SEX)
dim(nesarc.sub)
summary(nesarc.sub)
```

Running `summary` on the `nesarc.sub` reveals some non-obvious categories for the factors `CHECK321` and `S3AQ3B1`.  Reviewing the Code Book (HINT: Inside a pdf use `shift F` (Windows) or `command F` (Mac) then type the variable name inside the box to find the variable in the pdf), it is noted that a 9 is used to indicate missing values for `S3AQ3B1` and a 99 is used to indicate missing values for `S3AQ3C1`.

### Coding missing values

The variable `S3AQ3B1` uses a 9 to record `unknown` for smoking frequency and a 99 is used to record `unknown` for `S3AQ3C1`.


```{r label = "CodeMissing"}
nesarc.sub$S3AQ3B1[nesarc.sub$S3AQ3B1 == 9] <- NA
summary(nesarc.sub$S3AQ3B1)  # Note that 9 still appears
nesarc.sub$S3AQ3B1 <- factor(nesarc.sub$S3AQ3B1)[, drop = TRUE]
summary(nesarc.sub$S3AQ3B1)  # Unused level no longer appears
nesarc.sub$S3AQ3C1[nesarc.sub$S3AQ3C1 == 99] <- NA
summary(nesarc.sub$S3AQ3C1)
nesarc.sub$CHECK321[nesarc.sub$CHECK321 == 9] <- NA
nesarc.sub$CHECK321 <- factor(nesarc.sub$CHECK321)[ , drop = TRUE]
summary(nesarc.sub)
```

### Creating New Variables

To estimate the total number of cigarettes a subject smokes per month,  convert `S3AQ3B1` (a factor with 6 levels) to a numeric variable using `as.numeric`.  `DaysSmoke` estimates the days per month a subject smokes.  The variable `TotalCigsSmoked` estimates the monthly number of cigarettes a subject smokes per month by multiplying `DaysSmoke` times `S3AQ3C1` (the usual quantity smoked per day).  The numeric variable `TotalCigsSmoked` is converted into a factor with roughly equivalent numbers stored in each level of the factor `CigsSmokedFac` using the `cut` function.

```{r}
nesarc.sub$DaysSmoke <- as.numeric(nesarc.sub$S3AQ3B1)
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 1] <- 30
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 2] <- 4*5.5
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 3] <- 4*3.5
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 4] <- 4*1.5
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 5] <- 2.5
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 6] <- 1
# Using dplyr again
nesarc.sub <- nesarc.sub %>% 
  mutate(TotalCigsSmoked = DaysSmoke*S3AQ3C1)
proportions <- quantile(nesarc.sub$TotalCigsSmoked, na.rm = TRUE)
proportions
nesarc.sub$CigsSmokedFac <- cut(nesarc.sub$TotalCigsSmoked, breaks = proportions, include.lowest = TRUE)
head(nesarc.sub)
```

### Labeling Variables

In the next section of code, variables are labeled and levels of factors are given informative labels.  The order of the levels is also rearranged for the variables `S3AQ3B1`, `TAB12MDX`, and `SEX`. 

```{r}
library(Hmisc)
label(nesarc.sub$TAB12MDX) <- "Tobacco Dependence past 12 Months"
label(nesarc.sub$CHECK321) <- "Smoked Cigarettes in the Past 12 Months"
label(nesarc.sub$S3AQ3B1) <- "Usual Smoking Frequency"
label(nesarc.sub$S3AQ3C1) <- "Usual Smoking Quantity"
nesarc.sub$MAJORDEPLIFE <- factor(nesarc.sub$MAJORDEPLIFE, 
                              labels = c("No Depression", "Yes Depression"))
nesarc.sub$S3AQ3B1 <- factor(nesarc.sub$S3AQ3B1, 
                         labels = c("Every Day", "5 to 6 Days/week", 
                                    "3 to 4 Days/week",   "1 to 2 Days/week", "2 to 3 Days/month", 
                                    "Once a month or less"))
nesarc.sub$S3AQ3B1 <- factor(nesarc.sub$S3AQ3B1, 
                         levels = c("Once a month or less", "2 to 3 Days/month", "1 to 2 Days/week",  "3 to 4 Days/week", "5 to 6 Days/week", "Every Day"))
nesarc.sub$TAB12MDX <- factor(nesarc.sub$TAB12MDX, 
                         labels = c("No Nicotine Dependence", "Nicotine Dependence"))
nesarc.sub$TAB12MDX <- factor(nesarc.sub$TAB12MDX, 
                         levels = c("Nicotine Dependence", "No Nicotine Dependence"))
nesarc.sub$ETHRACE2A <- factor(nesarc.sub$ETHRACE2A, labels = c("Caucasian", "African American", "Native American", "Asian", "Hispanic"))
nesarc.sub$SEX <- factor(nesarc.sub$SEX, labels = c("Male", "Female"))
table(nesarc.sub$SEX)
nesarc.sub$SEX <- factor(nesarc.sub$SEX, levels = c("Female", "Male"))
summary(nesarc.sub)
```

### Renaming Variables

```{r}
nesarc.sub <- nesarc.sub %>% 
  rename(Depression = MAJORDEPLIFE, TobaccoDependence = TAB12MDX, SmokingFreq = S3AQ3B1, DailyCigsSmoked = S3AQ3C1, Ethnicity = ETHRACE2A, Sex = SEX, Age = AGE)
head(nesarc.sub)
```

### Creating Tables

Basic tables can be created using the functions `table` or `xtabs`. Frequency tables are created for the variables `TobaccoDependence`,  `CigsSmokedFac`, and `SmokingFreq`.

```{r}
T1 <- xtabs(~TobaccoDependence, data = nesarc.sub)
T2 <- xtabs(~CigsSmokedFac, data = nesarc.sub)
T3 <- xtabs(~SmokingFreq, data = nesarc.sub)
table(nesarc.sub$TobaccoDependence)
T1
T2
T3
```

In the data frame `nesarc.sub`, there are `r T1[1]` nicotine dependent subjects and `r T1[2]` subjects that are not nicotine dependent.  A small number of smokers (`r T2[4]`) smoke over 600 cigarettes per month. Most of the subjects in `nesarc.sub` are daily smokers (`r T3[6]`) with the remainder distributed uniformly across the first five levels of `SmokingFreq`.

### Graphing Frequency Tables

The barplots are all created with the package `ggplot2`.  The barplots start with the defaults for the `geom_bar` and add more detail to the plot with each graph.

```{r}
library(ggplot2)
g1 <- ggplot(data = nesarc.sub, aes(x = TobaccoDependence)) + 
  geom_bar() 
g1
g2 <- ggplot(data = na.omit(nesarc.sub), aes(x = CigsSmokedFac)) +
  geom_bar(fill = "brown")
g2
g3 <- ggplot(data = na.omit(nesarc.sub), aes(SmokingFreq)) +
  geom_bar(fill = "gray") +
  labs(x = "", y = "Total Number", title = "Smoking Frequency for Young Adults") +
  theme_bw() + 
  theme(axis.text.x  = element_text(angle = 75, vjust = 0.5)) 
g3
```

### Some Statistics (We have not discussed this yet)

```{r results='asis', echo = TRUE}
library(moonBook)
out <- mytable(Depression + TobaccoDependence ~  Age + SmokingFreq + DailyCigsSmoked + Ethnicity + Sex +DaysSmoke + TotalCigsSmoked + CigsSmokedFac , data = nesarc.sub)
myhtml(out)
```