---
title: "Data Management Example"
author: ""
date: "Spring 2015"
output: html_document
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, fig.align = "center")
library(ggplot2)
library(descr)
```

Three categorical variables relevant to *student health* are `H1SU1-H1SU3` from the `AddHealth` data set.  The `AddHealth` data set is loaded from the `PDS` package.  Using the `dplyr` package, a new data frame `NDF` containing only three variables after renaming `H1SU1-H1SU3` to `ConsideredSuicide`, `TimesAttemptedSuicide`, and `TreatedByDoctor` is created.


```{r}
library(PDS)
library(dplyr)
NDF <- AddHealth %>%
  rename(ConsideredSuicide = H1SU1, TimesAttemptedSuicide = H1SU2, TreatedByDoctor = H1SU3) %>%
  select(ConsideredSuicide, TimesAttemptedSuicide, TreatedByDoctor)
head(NDF)
```

Responses of `9` for question `ConsideredSuicide` are recoded to `NA`. 

```{r}
xtabs(~ConsideredSuicide, data = NDF)
NDF$ConsideredSuicide[NDF$ConsideredSuicide == 9] <- NA
xtabs(~ConsideredSuicide, data = NDF)
NDF$ConsideredSuicide <- factor(NDF$ConsideredSuicide, labels = c("No", "Yes", "Refused", "Don't Know"))[, drop = TRUE]
xtabs(~ConsideredSuicide, data = NDF)
summary(NDF$ConsideredSuicide)
freq(NDF$ConsideredSuicide)
freq(NDF$ConsideredSuicide, main = "Did you ever seriously think about committing suicide?")
ggplot(data = NDF, aes(x = ConsideredSuicide, fill = ConsideredSuicide)) + 
  geom_bar() + 
  labs(title = "Did you ever seriously think about committing suicide?", x = "") +
  theme_bw() +
  guides(fill = guide_legend(title = "Have You\nConsidered\nSuicide?"))
```


Changing the $y$-axis to percent:

```{r}
ggplot(data = NDF, aes(x = ConsideredSuicide, y = (..count..)/sum(..count..), fill = ConsideredSuicide)) + 
  geom_bar() + 
  labs(title = "Did you ever seriously think about committing suicide?", x = "") +
  theme_bw() +
  guides(fill = guide_legend(title = "Have You\nConsidered\nSuicide?")) +
  labs(y = "percent")
```

Another approach to graph the categorical variables is to use the base function `barplot()`.

```{r}
barplot(xtabs(~ConsideredSuicide, data = NDF), col = c("green", "red", "orange", "brown"))
barplot(prop.table(xtabs(~ConsideredSuicide, data = NDF)), col = c("green", "red", "orange", "brown"))
```

Question `TimesAttemptedSuicide` asks the 821 respondents from `ConsideredSuicide` "how many times did you attempt suicide?"  Responses of 7 are recoded to `NA`.

```{r}
xtabs(~TimesAttemptedSuicide, data = NDF)
NDF$TimesAttemptedSuicide[NDF$TimesAttemptedSuicide==7] <- NA
xtabs(~TimesAttemptedSuicide, data = NDF)
NDF$TimesAttemptedSuicide <- factor(NDF$TimesAttemptedSuicide)[, drop = TRUE]
xtabs(~TimesAttemptedSuicide, data = NDF)
NDF$TimesAttemptedSuicide <- factor(NDF$TimesAttemptedSuicide, 
                          labels = c("0 times", "1 time", "2 or 3 times", 
                                     "4 or 5 times", ">= 6 times", "refused", 
                                     "don't know"))
```



```{r}
summary(NDF$TimesAttemptedSuicide)
xtabs(~TimesAttemptedSuicide, data = NDF)
NAS <- sum(xtabs(~TimesAttemptedSuicide, data = NDF)[2:5])  # number that attempted suicide
NAS
freq(NDF$TimesAttemptedSuicide, main = "How many times did you actually attempt suicide?")
ggplot(data = NDF, aes(x = TimesAttemptedSuicide)) + 
  geom_bar(fill = "maroon") + 
  labs(title = "How many times did you actually attempt suicide?", x= "") +
  theme_bw() 
ggplot(data = na.omit(NDF[ , "TimesAttemptedSuicide", drop = FALSE]), aes(x = TimesAttemptedSuicide)) + 
  geom_bar(fill = "plum") + 
  labs(title = "How many times did you actually attempt suicide?", x = "") +
  theme_bw() 
ggplot(data = na.omit(NDF[ , "TimesAttemptedSuicide", drop = FALSE]), 
       aes(x = TimesAttemptedSuicide, fill = TimesAttemptedSuicide)) + 
  geom_bar() + 
  labs(title = "How many times did you actually attempt suicide?", x = "") +
  theme_bw() +
  guides(fill = guide_legend(title = "Number of Times\nSuicide Attempted"))
```

Question `TreatedByDoctor` asks of the `r NAS` that attempted suicide "Did you require treatment by either a nurse or doctor?"  A response of 7 should be recoded to `NA`.

```{r}
NDF$TreatedByDoctor[NDF$TreatedByDoctor == 7] <- NA
NDF$TreatedByDoctor <- factor(NDF$TreatedByDoctor, labels = c("No", "Yes"))[, drop = TRUE]
xtabs(~TreatedByDoctor, data = NDF)
freq(NDF$TreatedByDoctor, main = "Did you require treatment by either a nurse or doctor?")
ggplot(data = na.omit(NDF[ , "TreatedByDoctor", drop = FALSE]), aes(x = TreatedByDoctor, fill = TreatedByDoctor)) + 
  geom_bar() + 
  labs(title = "Did you require treatment by either a nurse or doctor?", x = "") +
  theme_bw() +
  guides(fill = guide_legend(title = "Treated by Doctor\n or Nurse"))
```