---
title: "NESARC Commands"
author: "Alan T. Arnholt"
date: "February 13, 2015"
output: html_document
---

```{r, label = "Setup", echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, fig.align = "center")
```

## Data Management

Start by loading the `PDS` package.  If you need directions for installing the `PDS` package, see [https://github.com/alanarnholt/PDS](https://github.com/alanarnholt/PDS).

```{r}
library(PDS)
```

Create a subset of people 25 or younger who smoked in the last 12 months.  To do this, one can use the `dplyr` package. Note that `CHECK321 == 1` is used to see if subject has smoked in the past 12 months and `!is.na()` is used to make sure the subset does not contain `NA`s for the chosen variables.

```{r}
library(dplyr)
nesarc.sub <- NESARC %>% 
  filter(!is.na(CHECK321) & !is.na(AGE) & CHECK321 ==1 & AGE <= 25)
dim(nesarc.sub)
```

The `nesarc.sub` data frame contains `r dim(nesarc.sub)[2]` variables.  Use the verb `select` from the `dplyr` package to select the variables `CHECK321`, `AGE` (age of subject), `TAB12MDX` (tobaco dependence past 12 months), `S3AQ3B1` (usual smoking quantity), `ETHRACE2A` (ethnicity), `SEX` (gender), and `S3AQ3C1` (usual smoking frequency).  

```{r}
nesarc.sub <- nesarc.sub %>% 
  select(CHECK321, AGE, TAB12MDX, S3AQ3B1, S3AQ3C1, ETHRACE2A, SEX)
dim(nesarc.sub)
summary(nesarc.sub)
```

Running `summary` on the `nesarc.sub` reveals some non-obvious categories for the factors (`CHECK321`, `TAB12MDX`, and `S3AQ3B1`).  Reviewing the Code Book (HINT: Inside a pdf use `shift F` (Windows) or `command F` (Mac) then type the variable name inside the box to find the variable in the pdf), it is noted that a 9 is used to indicate missing values for `S3AQ3B1` and a 99 is used to indicate missing values for `S3AQ3C1`.

### Coding missing values

The variable `S3AQ3B1` uses a 9 to record `unknown` for smoking frequency and a 99 is used to record `unknown` for `S3AQ3C1`.


```{r label = "CodeMissing"}
nesarc.sub$S3AQ3B1[nesarc.sub$S3AQ3B1 == 9] <- NA
summary(nesarc.sub$S3AQ3B1)  # Note that 9 still appears
nesarc.sub$S3AQ3B1 <- factor(nesarc.sub$S3AQ3B1)[, drop = TRUE]
summary(nesarc.sub$S3AQ3B1)  # Unused level no longer appears
nesarc.sub$S3AQ3C1[nesarc.sub$S3AQ3C1 == 99] <- NA
summary(nesarc.sub$S3AQ3C1)
summary(nesarc.sub)
```

### Creating New Variables

The goal now is to estimate the total number of cigarettes a subject smokes per month.  Using the `S3AQ3B1` variable, convert `S3AQ3B1` (a factor with 6 levels) to a numeric variable using `as.numeric`.  The new variable `DaysSmoke` estimates the days per month a subject smokes.  The variable `TotalCigsSmoked` estimates the monthly number of cigarettes a subject smokes per month by multiplying `DaysSmoke` times `S3AQ3C1` (the usual quantity smoked in a given day).

```{r}
nesarc.sub$DaysSmoke <- as.numeric(nesarc.sub$S3AQ3B1)
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 1] <- 30
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 2] <- 4*5.5
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 3] <- 4*3.5
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 4] <- 4*1.5
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 5] <- 2.5
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 6] <- 1
# Using dplyr again
nesarc.sub <- nesarc.sub %>% 
  mutate(TotalCigsSmoked = DaysSmoke*S3AQ3C1)
proportions <- quantile(nesarc.sub$TotalCigsSmoked, na.rm = TRUE)
proportions
nesarc.sub$CigsSmokedFac <- cut(nesarc.sub$TotalCigsSmoked, breaks = proportions, include.lowest = TRUE)
head(nesarc.sub)
```

### Labeling Variables

```{r}
library(Hmisc)
label(nesarc.sub$TAB12MDX) <- "Tobacco Dependence past 12 Months"
label(nesarc.sub$CHECK321) <- "Smoked Cigarettes in the Past 12 Months"
label(nesarc.sub$S3AQ3B1) <- "Usual Smoking Frequency"
label(nesarc.sub$S3AQ3C1) <- "Usual Smoking Quantity"
nesarc.sub$S3AQ3B1 <- factor(nesarc.sub$S3AQ3B1, 
                         labels = c("Every Day", "5 to 6 Days/week", 
                                    "3 to 4 Days/week",   "1 to 2 Days/week", "2 to 3 Days/month", 
                                    "Once a month or less"))
nesarc.sub$S3AQ3B1 <- factor(nesarc.sub$S3AQ3B1, 
                         levels = c("Once a month or less", "2 to 3 Days/month", "1 to 2 Days/week",  "3 to 4 Days/week", "5 to 6 Days/week", "Every Day"))
nesarc.sub$TAB12MDX <- factor(nesarc.sub$TAB12MDX, 
                         labels = c("No Nicotine Dependence", "Nicotine Dependence"))
nesarc.sub$TAB12MDX <- factor(nesarc.sub$TAB12MDX, 
                         levels = c("Nicotine Dependence", "No Nicotine Dependence"))

table(nesarc.sub$TAB12MDX)
nesarc.sub$ETHRACE2A <- factor(nesarc.sub$ETHRACE2A, labels = c("Caucasian", "African American", "Native American", "Asian", "Hispanic"))
nesarc.sub$SEX <- factor(nesarc.sub$SEX, labels = c("Male", "Female"))
table(nesarc.sub$SEX)
nesarc.sub$SEX <- factor(nesarc.sub$SEX, levels = c("Female", "Male"))
table(nesarc.sub$SEX)
```

### Renaming Variables

```{r}
nesarc.sub <- nesarc.sub %>% 
  rename(TobacoDependence = TAB12MDX, SmokingFreq = S3AQ3B1, DailyCigsSmoked = S3AQ3C1, Ethnicity = ETHRACE2A, Sex = SEX, Age = AGE)
head(nesarc.sub)
```


### Representing Categorical Data

```{r}
T1 <- xtabs(~TobacoDependence, data = nesarc.sub)
T1
barplot(T1)
T2 <- xtabs(~TobacoDependence + SmokingFreq, data = nesarc.sub)
T2
PT2 <- prop.table(T2, 2)
PT2
barplot(PT2)
```

```{r fig.width = 10, fig.height = 10}
library(ggplot2)
ggplot(data = na.omit(nesarc.sub), aes(x = SmokingFreq, fill = TobacoDependence)) + 
  geom_bar(position = "fill") + 
  theme_bw()
```

The labels on the `x-axis` of the last graph are overlapping.

```{r fig.width = 10, fig.height = 10}
library(ggplot2)
ggplot(data = na.omit(nesarc.sub), aes(x = SmokingFreq, fill = TobacoDependence)) + 
  geom_bar(position = "fill") + 
  theme_bw() + 
  theme(axis.text.x  = element_text(angle = 45, vjust = 0.5)) +
  labs(x = "Smoking Frequency", y = "Percent")
```

Does the relationship hold for different genders?

```{r fig.width = 10, fig.height = 10}
library(ggplot2)
ggplot(data = na.omit(nesarc.sub), aes(x = SmokingFreq, fill = TobacoDependence)) + 
  geom_bar(position = "fill") + 
  theme_bw() + 
  theme(axis.text.x  = element_text(angle = 45, vjust = 0.5)) +
  labs(x = "Smoking Frequency", y = "Percent") + 
  facet_grid(Sex ~ .)
```


Does the relationship hold for different ethnicities?

```{r fig.width = 10, fig.height = 10}
library(ggplot2)
ggplot(data = na.omit(nesarc.sub), aes(x = SmokingFreq, fill = TobacoDependence)) + 
  geom_bar(position = "fill") + 
  theme_bw() + 
  theme(axis.text.x  = element_text(angle = 45, vjust = 0.5)) +
  labs(x = "Smoking Frequency", y = "Percent") + 
  facet_grid(Ethnicity ~ .)
```


```{r fig.width = 10, fig.height = 10}
library(ggplot2)
ggplot(data = na.omit(nesarc.sub), aes(x = CigsSmokedFac, fill = TobacoDependence)) + 
  geom_bar(position = "fill") + 
  theme_bw() + 
  theme(axis.text.x  = element_text(angle = 45, vjust = 0.5)) +
  labs(x = "Smoking Quantity", y = "Percent") 
```

```{r fig.width = 10, fig.height = 10}
library(ggplot2)
ggplot(data = na.omit(nesarc.sub), aes(x = CigsSmokedFac, fill = TobacoDependence)) + 
  geom_bar(position = "fill") + 
  theme_bw() + 
  theme(axis.text.x  = element_text(angle = 45, vjust = 0.5)) +
  labs(x = "Smoking Quantity", y = "Percent") + 
  facet_grid(Ethnicity ~ Sex)
```