---
title: "Displaying Categorical Data"
author: "R-Man"
date: "`r Sys.Date()`"
output: html_document
---

```{r, label = "SETUP", echo = FALSE, results= 'hide', message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(comment = NA, fig.show = 'as.is', prompt = TRUE, highlight = TRUE, tidy = FALSE, warning = FALSE, message = FALSE, tidy.opts=list(blank = TRUE, width.cutoff= 75))
```


## Who sailed on the Titanic?

The data set `TITANIC3` from the `PASWR2` package contains survival status of individual passengers on the Titanic but no information for the crew.  Start by creating simple univariate barcharts of the passengers.

```{r}
library(PASWR2)
library(ggplot2)
TITANIC3[1:5, 1:3]  # first five rows and three variables
# Create a factor out of the survived variable
TITANIC3$survived <- factor(TITANIC3$survived, labels = c("No", "Yes"))
TITANIC3[1:5, 1:3]  # first five rows and three variables
p1 <- ggplot(data = TITANIC3, aes(x = pclass)) + 
  geom_bar(fill = "purple") + 
  theme_bw() + 
  labs(x = "Ticket Class of Passenger", y = "", 
       title = "Your Informative Title\n Goes Here")
p1 
p2 <- ggplot(data = TITANIC3, aes(x = sex, fill = sex)) + 
  geom_bar() + 
  theme_bw() + 
  scale_fill_manual(values = c("pink", "blue"), guide = FALSE)
p2
p3 <- ggplot(data = TITANIC3, aes(x = sex, fill = survived)) + 
  geom_bar(position = "fill") + 
  theme_bw() + 
  scale_fill_manual(values = c("gray", "green")) + 
  labs(y = "Fraction")
p3
```

Let us arrange all three barplots on the same page.

```{r}
gridExtra::grid.arrange(p1, p2, p3, nrow = 1, widths = c(3, 2, 3))
```

Suppose we would like to look at the same graphs for the subset of passengers that were between ages 20 and 30 includsive.

```{r}
library(dplyr)
age2030 <- TITANIC3 %>% 
  filter(age >= 20, age <= 30)
p123 <- ggplot(data = age2030, aes(x = pclass)) + 
  geom_bar(fill = "purple") + 
  theme_bw()
p223 <- ggplot(data = age2030, aes(x = sex, fill = sex)) + 
  geom_bar() + 
  theme_bw() + 
  scale_fill_manual(values = c("pink", "blue"), guide = FALSE)
p323 <- ggplot(data = age2030, aes(x = sex, fill = survived)) + 
  geom_bar(position = "fill") + 
  theme_bw() + 
  scale_fill_manual(values = c("gray", "green")) + 
  labs(y = "Fraction")
gridExtra::grid.arrange(p123, p223, p323, nrow = 1, widths = c(3, 2, 3))
```

Next we will create a factor variable for the various values in `age`

```{r}
TITANIC3$agef <- cut(TITANIC3$age, breaks = c(0, 15, 30, 60, 80))
summary(TITANIC3$agef)
pf1 <- ggplot(data = TITANIC3, aes(x = pclass)) + 
  geom_bar(fill = "purple") + 
  theme_bw() + 
  facet_grid(agef ~ .)
pf2 <- ggplot(data = TITANIC3, aes(x = sex, fill = sex)) + 
  geom_bar() + 
  theme_bw() + 
  scale_fill_manual(values = c("pink", "blue"), guide = FALSE) + 
  facet_grid(agef ~ .)
pf3 <- ggplot(data = TITANIC3, aes(x = sex, fill = survived)) + 
  geom_bar(position = "fill") + 
  theme_bw() + 
  scale_fill_manual(values = c("gray", "green")) + 
  labs(y = "Fraction") + 
  facet_grid(agef ~ .)
gridExtra::grid.arrange(pf1, pf2, pf3, nrow = 1, main = "Your Informative Title\n Goes Here", widths = c(2, 2, 3))
```

## Studying Multivariate Categorical Data

```{r}
ggplot(data = TITANIC3, aes(x = pclass, fill = survived)) + 
  geom_bar(position = "dodge") + 
  facet_grid(agef~ sex) + 
  theme_bw()
```

## Doubledecker Plots

```{r}
vcd::doubledecker(survived ~ sex, data = TITANIC3)
vcd::doubledecker(survived ~ sex + pclass, data = TITANIC3)
```

## Mosaic Plots

```{r}
vcd::mosaic(survived ~ sex + pclass, data = TITANIC3)
vcd::mosaic(~ sex + survived + pclass, data = TITANIC3, shade = TRUE)
```