---
title: "More Data Management and Graphs"
author: ""
date: "Spring 2015"
output: html_document
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, fig.align = "center")
library(dplyr)
library(ggplot2)
```


```{r label = "Missing"}
library(PDS)
NESARC$S3AQ3B1[NESARC$S3AQ3B1 == 9] <- NA
summary(NESARC$S3AQ3B1)  # Note that 9 still appears
NESARC$S3AQ3B1 <- factor(NESARC$S3AQ3B1)[, drop = TRUE]
summary(NESARC$S3AQ3B1)  # Unused level no longer appears
NESARC$S3AQ3B1 <- factor(NESARC$S3AQ3B1, 
                         labels = c("Every Day", "5 to 6 Days/week", 
                                    "3 to 4 Days/week", "1 to 2 Days/week", 
                                    "2 to 3 Days/month", "Once a month or less"))
summary(NESARC$S3AQ3B1)
xtabs(~S3AQ3B1, data = NESARC) # Note how the NA's are not printed
```

Subsetting the data to individuals who have smoke over 100 cigarettes (`S3AQ1A ==1`), have smoked cigarettes in the past 12 months (`CHECK321 == 1`), indicated they are daily smokers (`S3AQ3B1 == "Every Day"`), and are between the ages of 18 and 25 (`AGE <= 25` & `AGE >= 18`).

```{r, label = "Subsetting"}
NESARCsub1 <- NESARC %>%
  filter(S3AQ1A == 1 & CHECK321 == 1 & S3AQ3B1 == "Every Day" & AGE <= 25 & AGE >= 18)
dim(NESARCsub1) 
```

Creating a 5 level factor out of the numeric variable `S3AQ3C1` which records the number of cigarettes smoked per day.
```{r}
NESARCsub1$S3AQ3C1[NESARCsub1$S3AQ3C1 == 99] <- NA
summary(NESARCsub1$S3AQ3C1)
NESARCsub1$S3AQ3C1fac <- cut(NESARCsub1$S3AQ3C1, breaks = c(1, 5, 10, 15, 20, 98))
summary(NESARCsub1$S3AQ3C1fac)
NESARCsub1$SEX <- factor(NESARCsub1$SEX, labels = c("Male", "Female")) 
NESARCsub1$ETHRACE2A <- factor(NESARCsub1$ETHRACE2A, labels = c("Caucasian", "African American", "American Indian", "Asian", "Hispanic")) 
```

```{r}
C1 <- with(data = NESARCsub1, table(S3AQ3C1fac))
C1
PC1 <- prop.table(C1)
PC1
T1 <- with(data = NESARCsub1, ftable(SEX, S3AQ3C1fac, ETHRACE2A))
T1
prop.table(T1, 2)
apply(prop.table(T1, 2), 2, sum)
T2 <- with(data = NESARCsub1, ftable(ETHRACE2A, SEX, S3AQ3C1fac))
T2
PT3 <- prop.table(T2, 1)
PT3
```

## Creating Bar Charts

```{r}
# No frills (base) bar chart
barplot(C1)
# No frills (ggplot2) bar chart
ggplot(data = NESARCsub1, aes(x = S3AQ3C1fac)) + geom_bar()
# Similar to book (base)
barplot(PC1, col = "blue", xlab = "Number of cigarettes smoked/day", ylab = "Percent")
# Similar to book (ggplot2)
ggplot(data = na.omit(NESARCsub1[ ,"S3AQ3C1fac", drop = FALSE]), aes(x = S3AQ3C1fac, y = (..count..)/(sum(..count..)))) + 
  geom_bar(fill = "blue") +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent")
```

# Explain what is happening in this graph

```{r}
ggplot(data = na.omit(NESARCsub1[ , c("S3AQ3C1fac", "SEX", "ETHRACE2A"), drop = FALSE]), aes(x = S3AQ3C1fac, y = (..count..)/(sum(..count..)), fill = SEX)) + 
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent") + 
  facet_grid(ETHRACE2A ~ .)
```

# Caucasians now (explain what is going on)
```{r}
ggplot(data = subset(na.omit(NESARCsub1[ , c("S3AQ3C1fac", "SEX", "ETHRACE2A"), drop = FALSE]), ETHRACE2A == "Caucasian"), aes(x = S3AQ3C1fac, y = (..count..)/(sum(..count..)), fill = SEX)) + 
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent")
# Comapre to
ggplot(data = subset(na.omit(NESARCsub1[ , c("S3AQ3C1fac", "SEX", "ETHRACE2A"), drop = FALSE]), ETHRACE2A == "Caucasian"), aes(x = S3AQ3C1fac, y = ..density.., fill = SEX, group = SEX)) + 
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent")
```

# African Americans

```{r}
ggplot(data = subset(na.omit(NESARCsub1[ , c("S3AQ3C1fac", "SEX", "ETHRACE2A"), drop = FALSE]), ETHRACE2A == "African American"), aes(x = S3AQ3C1fac, y = (..count..)/(sum(..count..)), fill = SEX)) + 
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent")
# Compare to
ggplot(data = subset(na.omit(NESARCsub1[ , c("S3AQ3C1fac", "SEX", "ETHRACE2A"), drop = FALSE]), ETHRACE2A == "African American"), aes(x = S3AQ3C1fac, y = ..density.., fill = SEX, group = SEX)) + 
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent")
```

# American Indians

```{r}
ggplot(data = subset(na.omit(NESARCsub1[ , c("S3AQ3C1fac", "SEX", "ETHRACE2A"), drop = FALSE]), ETHRACE2A == "American Indian"), aes(x = S3AQ3C1fac, y = (..count..)/(sum(..count..)), fill = SEX)) + 
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent")
## Compare to
ggplot(data = subset(na.omit(NESARCsub1[ , c("S3AQ3C1fac", "SEX", "ETHRACE2A"), drop = FALSE]), ETHRACE2A == "American Indian"), aes(x = S3AQ3C1fac, y = ..density.., fill = SEX, group = SEX)) + 
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent")
```

# Asian

```{r}
ggplot(data = subset(na.omit(NESARCsub1[ , c("S3AQ3C1fac", "SEX", "ETHRACE2A"), drop = FALSE]), ETHRACE2A == "Asian"), aes(x = S3AQ3C1fac, y = (..count..)/(sum(..count..)), fill = SEX)) + 
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent")
## Compare to
ggplot(data = subset(na.omit(NESARCsub1[ , c("S3AQ3C1fac", "SEX", "ETHRACE2A"), drop = FALSE]), ETHRACE2A == "Asian"), aes(x = S3AQ3C1fac, y = ..density.., fill = SEX, group = SEX)) + 
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent")
```

# Hispanic

```{r}
ggplot(data = subset(na.omit(NESARCsub1[ , c("S3AQ3C1fac", "SEX", "ETHRACE2A"), drop = FALSE]), ETHRACE2A == "Hispanic"), aes(x = S3AQ3C1fac, y = (..count..)/(sum(..count..)), fill = SEX)) + 
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent")
###
ggplot(data = subset(na.omit(NESARCsub1[ , c("S3AQ3C1fac", "SEX", "ETHRACE2A"), drop = FALSE]), ETHRACE2A == "Hispanic"), aes(x = S3AQ3C1fac, y = ..density.., fill = SEX, group = SEX)) + 
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent")
```

# Is this the graph we want? 

```{r fig.width = 8, fig.height = 12}
PT3
DFM <- na.omit(NESARCsub1[ , c("S3AQ3C1fac", "SEX", "ETHRACE2A")])
dim(DFM)
ggplot(data = DFM, aes(x = S3AQ3C1fac)) +  
  geom_bar(position = "dodge", aes(y = ..density.., group = SEX, fill = SEX)) + 
  facet_grid(ETHRACE2A ~ .) +
  theme_bw() + 
  labs(x = "Number of cigarettes smoked/day", y = "Percent") + 
  scale_fill_manual(values = c("blue", "pink")) +
  guides(fill = guide_legend(title = "Gender")) 
```


