---
title: "More Graphs"
author: "Alan Arnholt"
date: "Spring 2015"
output: html_document
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, fig.align = "center")
```

I am trying to create side-by-side comparisons using `geom_bar` while conditioning on one and two factors.  Below I simulate the data frame `DF`.

```{r}
library(ggplot2)
set.seed(13)
fac1 <- factor(sample(letters[1:5], 500, replace = TRUE))
fac2 <- factor(sample(LETTERS[1:3], 500, replace = TRUE))
sex <- factor(sample(c("Male", "Female"), 500, replace = TRUE))
DF <- data.frame(fac1, fac2, sex)
rm(fac1, fac2, sex)
head(DF)
```

When there is a single factor, I get what I would expect with the code below.

```{r}
T1 <- xtabs(~ fac1, data = DF)
PT1 <- prop.table(T1)
PT1
ggplot(data = DF, aes(x = fac1)) + 
  geom_bar(aes(y = ..count../sum(..count..)), fill = "lightblue") +
  geom_hline(y = PT1[1], linetype = "dashed") + 
  theme_bw() + 
  labs(y = "Percent")
```

I want to create barplots on the first factor (`fac1`) while conditioning on the second factor (`fac2`). Specifically, I expected the height of the bars for each level of `fac2` to sum to 1 for each of the three `facets`.  

```{r}
T2 <- ftable(DF$fac2, DF$fac1)
PT2 <- prop.table(T2, 1)
PT2
apply(PT2, 1, sum) # quick check that sum is 1
#
library(dplyr)
NT <- DF %>%
  count(fac1, fac2)
NT
perc <- matrix(PT2, nrow = 15)
SN <- cbind(NT, perc)
rm(perc)
SN
ggplot(data = SN, aes(x = fac1)) + 
  geom_bar(stat = "identity", aes(y = perc), fill = "lightblue") + 
  facet_grid(fac2~.) + 
  theme_bw() +
  labs(y = "Percent")
# Similar type graph...but not quite as it is harder to read
ggplot(DF, aes(x = fac2, fill = fac1)) +
   theme_bw() +
   geom_bar(position = "fill") +
   ylab("Proportion")
```

I would like to create barplots conditioning on both `fac2` and `sex` such that the height of the bars would agree with the values in `PT3`.  The heights of the bars for `Male` and `Female` in each `facet` sum to 1.

```{r}
T3 <- ftable(DF$fac2, DF$sex, DF$fac1)
PT3 <- prop.table(T3, 1)
PT3
apply(PT3, 1, sum) # quick check that all sums to 1
NS <- DF %>% 
  count(fac1, fac2, sex)
NS
perc2 <- matrix(PT3, nrow=30)
NDF <- cbind(NS, perc2)
rm(perc2)
head(NDF)
### 
ggplot(data = NDF, aes(x = fac1)) +
  facet_grid(fac2 ~.) +
  geom_bar(position = "dodge", stat = "identity", aes(y = perc2, fill = sex)) +
  theme_bw() +
  labs(y = "Percent") +
  scale_fill_manual(values = c("blue", "pink"))  
```