---
title: "Reading Data"
author: "R-Man"
date: "`r Sys.Date()`"
output: html_document
---

Download the `csv` files you need from [http://nces.ed.gov/ipeds/datacenter/DataFiles.aspx](http://nces.ed.gov/ipeds/datacenter/DataFiles.aspx).  As an example, I have downloaded the files `ic2013_rv.csv` and `hd2013.csv`.   To read a file, use the `read.csv` function.  Note that these `csv` files use a `.` to denote a missing value. 

```{r}
IC2013 <- read.csv(file = "ic2013_rv.csv", na.strings = ".")
IC2013[1:6, 1:8]
IC2013 <- dplyr::tbl_df(IC2013)
IC2013
dim(IC2013)
HD2013 <- read.csv(file = "hd2013.csv", na.strings = ".")
dim(HD2013)
HD2013 <- dplyr::tbl_df(HD2013)
HD2013
HD2013[1:6, 1:2]
```

To merge the two files, use the `full_join` function from `dplyr`.

```{r}
MERGED <- dplyr::full_join(IC2013, HD2013, by = "UNITID")
dim(MERGED)
MERGED[1:6, 200:205]
MERGED
```

## More

```{r}
EF2013A <- read.csv(file = "ef2013a.csv", na.strings = ".")
head(EF2013A)
library(tidyr)
newef <- EF2013A %>% 
  gather(RaceEthnic, Enrolled, EFTOTLT:EFNRALW)
head(newef)
dim(newef)
tail(newef, n = 250)
```


Consider reading in an SPSS file.  Consider the file (`kidshtwt.sav`) containing the kids and their heights at one year of age (`HT1`) and at two years of age (`HT2`) as well as weights at one year of age (`WT1`) and at two years of age (`WT2`).

```{r}
library(foreign)
ND <- read.spss(file = "kidshtwt.sav", to.data.frame = TRUE)
ND
boxplot(ND$HT1, ND$HT2)
library(tidyr)
nd <- ND %>% 
gather(Age, Height, HT1:HT2) %>% 
  gather(Age2, Weight, WT1:WT2)
head(nd)
boxplot(Height ~ Age, data = nd)
library(ggplot2)
ggplot(data = nd, aes(x = Age, y = Height)) + geom_boxplot()
```
