---
title: "Hypotheses Testing"
author: "Alan Arnholt"
date: "`r Sys.Date()`"
output: html_document
params:
  seed: 154
  alt: "two.sided"
  alpha: 0.05
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, fig.align = "center")
```


Generating some random data:

```{r}
set.seed(params$seed)
score <- c(rnorm(25, 95, 15), rnorm(25, 105, 15))
school <- gl(n = 2, k = 25, labels = c("Blowing Rock", "Hardin Park"))
DF <- data.frame(score = score, school = school)
```

Test whether the mean IQ score is the same for Blowing Rock and Hardin Park versus
whether the mean IQ score for Blowing Rock is $`r ifelse(params$alt == "two.sided", "\\neq", ifelse(params$alt == "greater", ">", "<"))`$ the mean IQ score for Hardin Park.

$$H_0: \mu_{Blowing Rock} = \mu_{Hardin Park}$$ versus $$H_1: \mu_{Blowing Rock} `r ifelse(params$alt == "two.sided", "\\neq", ifelse(params$alt == "greater", ">", "<"))` \mu_{Hardin Park}$$

Graphs first:

```{r}
library(ggplot2)
ggplot(data = DF, aes(x = school, y = score)) +
  geom_boxplot() + 
  theme_bw()
```

Test next:

```{r}
TR <- t.test(score ~ school, data = DF, var.equal = TRUE, alt = params$alt)
TR
```

* Statistical conclusion: `r ifelse(TR$p.value < params$alpha, "reject the null hypothesis.", "fail to reject the null hypothesis.")` 
* English conclusion: `r ifelse(TR$p.value < params$alpha, paste("evidence suggests the mean IQ scores for Blowing Rock students is", ifelse(params$alt == "two.sided", " not equal to ", ifelse(params$alt == "greater", " greater than ", " less than ")), "the mean IQ scores for Hardin Park students.", sep = ""), paste("insufficient evidence to suggests the mean IQ scores for Blowing Rock students is", ifelse(params$alt == "two.sided", " not equal to ", ifelse(params$alt == "greater", " greater than ", " less than ")), "the mean IQ scores for Hardin Park students.", sep = ""))`



The ANOVA test is equivalent to the `t.test` when the alternative hypothesis is `two.sided` and `var.equal = TRUE`:

```{r}
mod <- aov(score ~ school, data = DF)
AR <- summary(mod)
AR
```

* Statistical conclusion: `r ifelse(AR[[1]][1, 5] < params$alpha, "reject the null hypothesis.", "fail to reject the null hypothesis.")`
* English conclusion: `r ifelse(AR[[1]][1, 5] < params$alpha, "evidence suggests an association between average IQ and school district exists.", "not enough evidence to suggest an association between average IQ and school district exists.")`


Generating `DF2`:

```{r}
set.seed(params$seed)
score <- c(rnorm(25, 95, 15), rnorm(25, 100, 15), rnorm(25, 105, 15))
school <- gl(n = 3, k = 25, labels = c("Two Rivers", "Blowing Rock", "Hardin Park"))
DF2 <- data.frame(score = score, school = school)
```

Graphs first:

```{r}
ggplot(data = DF2, aes(x = school, y = score)) + 
  geom_boxplot() + 
  theme_bw()
```

```{r}
mod <- aov(score ~ school, data = DF2)
AR <- summary(mod)
AR
```

* Statistical conclusion: `r ifelse(AR[[1]][1, 5] < params$alpha, "reject the null hypothesis.", "fail to reject the null hypothesis.")`
* English conclusion: `r ifelse(AR[[1]][1, 5] < params$alpha, "evidence suggests an association between average IQ and school district exists.", "not enough evidence to suggest an association between average IQ and school district exists.")`

Post Hoc Tests:

```{r}
TukeyHSD(mod)
opar <- par(no.readonly = TRUE)
par(mar = c(5.1, 11.1, 4.1, 2.1), las = 1) # Enlarge left margin
plot(TukeyHSD(mod))
par(opar)  # return to defaults
```