---
title: "Project Template"
author: "Alan T. Arnholt"
date: "March 30, 2015"
output: html_document
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, fig.align = "center")
```

### Purpose of Study: (Note: This is the last paragraph of the research proposal used to remind the reader of the research objectives.)

The present study will examine young adults from the National Epidemiologic Survey of Alcohol and Related Conditions (NESARC). The goals of the analysis will include 1) establishing the relationship between major depression and nicotine dependence; and 2) determining whether or not the relationship between major depression and nicotine dependence exists above and beyond smoking quantity. 

### Variables:

Variables from NESARC that will be used include: `MAJORDEPLIFE` (has subject experienced a major depression?), `CHECK321` (has subject smoked cigarettes in past 12 months?), `AGE` (age of subject), `TAB12MDX` (tobacco dependence past 12 months), `S3AQ3B1` (usual frequency when cigarettes smoked), `ETHRACE2A` (ethnicity of subject), `SEX` (gender of subject), and `S3AQ3C1` (usual smoking quantity of cigarettes when smoked).  ￼

### Data Management

First, the data is placed on the search path using the `PDS` package.
Next, a subset of people 25 or younger who smoked in the last 12 months is created using the `filter` function from `dplyr`. Note that `CHECK321 == 1` is used to see if subject has smoked in the past 12 months and `!is.na()` is used to make sure the subset does not contain `NA`s for the chosen variables.  Last, the variables of interest are selected and stored in the data frame `nesarc.sub` using the `select` function from the `dplyr` package.

```{r}
library(PDS)
library(dplyr)
nesarc.sub <- NESARC %>% 
  filter(!is.na(CHECK321) & !is.na(AGE) & CHECK321 ==1 & AGE <= 25)
dim(nesarc.sub)
nesarc.sub <- nesarc.sub %>% 
  select(MAJORDEPLIFE, CHECK321, AGE, TAB12MDX, S3AQ3B1, S3AQ3C1, ETHRACE2A, SEX)
dim(nesarc.sub)
summary(nesarc.sub)
```

Running `summary` on the `nesarc.sub` reveals some non-obvious categories for the factors `CHECK321` and `S3AQ3B1`.  Reviewing the Code Book (HINT: Inside a pdf use `shift F` (Windows) or `command F` (Mac) then type the variable name inside the box to find the variable in the pdf), it is noted that a 9 is used to indicate missing values for `S3AQ3B1` and a 99 is used to indicate missing values for `S3AQ3C1`.

### Coding missing values

The variable `S3AQ3B1` uses a 9 to record `unknown` for smoking frequency and a 99 is used to record `unknown` for `S3AQ3C1`.


```{r label = "CodeMissing"}
nesarc.sub$S3AQ3B1[nesarc.sub$S3AQ3B1 == 9] <- NA
summary(nesarc.sub$S3AQ3B1)  # Note that 9 still appears
nesarc.sub$S3AQ3B1 <- factor(nesarc.sub$S3AQ3B1)[, drop = TRUE]
summary(nesarc.sub$S3AQ3B1)  # Unused level no longer appears
nesarc.sub$S3AQ3C1[nesarc.sub$S3AQ3C1 == 99] <- NA
summary(nesarc.sub$S3AQ3C1)
nesarc.sub$CHECK321[nesarc.sub$CHECK321 == 9] <- NA
nesarc.sub$CHECK321 <- factor(nesarc.sub$CHECK321)[ , drop = TRUE]
summary(nesarc.sub)
```

### Creating New Variables

To estimate the total number of cigarettes a subject smokes per month,  convert `S3AQ3B1` (a factor with 6 levels) to a numeric variable using `as.numeric`.  `DaysSmoke` estimates the days per month a subject smokes.  The variable `TotalCigsSmoked` estimates the monthly number of cigarettes a subject smokes per month by multiplying `DaysSmoke` times `S3AQ3C1` (the usual quantity smoked per day).  The numeric variable `TotalCigsSmoked` is converted into a factor with roughly equivalent numbers stored in each level of the factor `CigsSmokedFac` using the `cut` function.

```{r}
nesarc.sub$DaysSmoke <- as.numeric(nesarc.sub$S3AQ3B1)
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 1] <- 30
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 2] <- 4*5.5
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 3] <- 4*3.5
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 4] <- 4*1.5
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 5] <- 2.5
nesarc.sub$DaysSmoke[nesarc.sub$DaysSmoke == 6] <- 1
# Using dplyr again
nesarc.sub <- nesarc.sub %>% 
  mutate(TotalCigsSmoked = DaysSmoke*S3AQ3C1)
proportions <- quantile(nesarc.sub$TotalCigsSmoked, na.rm = TRUE)
proportions
nesarc.sub$CigsSmokedFac <- cut(nesarc.sub$TotalCigsSmoked, breaks = proportions, include.lowest = TRUE)
head(nesarc.sub)
```

### Labeling Variables

In the next section of code, variables are labeled and levels of factors are given informative labels.  The order of the levels is also rearranged for the variables `S3AQ3B1`, `TAB12MDX`, and `SEX`. 

```{r}
library(Hmisc)
label(nesarc.sub$TAB12MDX) <- "Tobacco Dependence past 12 Months"
label(nesarc.sub$CHECK321) <- "Smoked Cigarettes in the Past 12 Months"
label(nesarc.sub$S3AQ3B1) <- "Usual Smoking Frequency"
label(nesarc.sub$S3AQ3C1) <- "Usual Smoking Quantity"
nesarc.sub$MAJORDEPLIFE <- factor(nesarc.sub$MAJORDEPLIFE, 
                              labels = c("No Depression", "Yes Depression"))
nesarc.sub$S3AQ3B1 <- factor(nesarc.sub$S3AQ3B1, 
                         labels = c("Every Day", "5 to 6 Days/week", 
                                    "3 to 4 Days/week",   "1 to 2 Days/week", 
                                    "2 to 3 Days/month", 
                                    "Once a month or less"))
nesarc.sub$S3AQ3B1 <- factor(nesarc.sub$S3AQ3B1, 
                         levels = c("Once a month or less", "2 to 3 Days/month", 
                                    "1 to 2 Days/week",  "3 to 4 Days/week", 
                                    "5 to 6 Days/week", "Every Day"))
nesarc.sub$TAB12MDX <- factor(nesarc.sub$TAB12MDX, 
                         labels = c("No Nicotine Dependence", "Nicotine Dependence"))
nesarc.sub$TAB12MDX <- factor(nesarc.sub$TAB12MDX, 
                         levels = c("Nicotine Dependence", "No Nicotine Dependence"))
nesarc.sub$ETHRACE2A <- factor(nesarc.sub$ETHRACE2A, 
                               labels = c("Caucasian", "African American", 
                                          "Native American", "Asian", "Hispanic"))
nesarc.sub$SEX <- factor(nesarc.sub$SEX, labels = c("Male", "Female"))
table(nesarc.sub$SEX)
nesarc.sub$SEX <- factor(nesarc.sub$SEX, levels = c("Female", "Male"))
summary(nesarc.sub)
```

### Renaming Variables

```{r}
nesarc.sub <- nesarc.sub %>% 
  rename(Depression = MAJORDEPLIFE, TobaccoDependence = TAB12MDX, 
         SmokingFreq = S3AQ3B1, DailyCigsSmoked = S3AQ3C1, Ethnicity = ETHRACE2A, 
         Sex = SEX, Age = AGE)
head(nesarc.sub)
```

### Creating Tables

Basic tables can be created using the functions `table` or `xtabs`. Frequency tables are created for the variables `TobaccoDependence`,  `CigsSmokedFac`, and `SmokingFreq`.

```{r}
T1 <- xtabs(~TobaccoDependence, data = nesarc.sub)
T2 <- xtabs(~CigsSmokedFac, data = nesarc.sub)
T3 <- xtabs(~SmokingFreq, data = nesarc.sub)
table(nesarc.sub$TobaccoDependence)
T1
T2
T3
```

In the data frame `nesarc.sub`, there are `r T1[1]` nicotine dependent subjects and `r T1[2]` subjects that are not nicotine dependent.  A small number of smokers (`r T2[4]`) smoke over 600 cigarettes per month. Most of the subjects in `nesarc.sub` are daily smokers (`r T3[6]`) with the remainder distributed uniformly across the first five levels of `SmokingFreq`.

### Graphing Frequency Tables

The barplots are all created with the package `ggplot2`.  The barplots start with the defaults for the `geom_bar` and add more detail to the plot with each graph.

```{r}
library(ggplot2)
g1 <- ggplot(data = nesarc.sub, aes(x = TobaccoDependence)) + 
  geom_bar() 
g1
g2 <- ggplot(data = na.omit(nesarc.sub), aes(x = CigsSmokedFac)) +
  geom_bar(fill = "brown")
g2
g3 <- ggplot(data = na.omit(nesarc.sub), aes(SmokingFreq)) +
  geom_bar(fill = "gray") +
  labs(x = "", y = "Total Number", title = "Smoking Frequency for Young Adults") +
  theme_bw() + 
  theme(axis.text.x  = element_text(angle = 75, vjust = 0.5)) 
g3
```

### Graphing Numeric Variables

One popular way to graph a continuous variable is to use a histogram.  `R` has the base function `hist` which can be used for this purpose.  We will also use the package `ggplot2` to create histograms.  We start with a basic histogram of the variable `TotalCigsSmoked`.

```{r}
hist(nesarc.sub$TotalCigsSmoked)
```

Experiment with the code in the previous code chunk to change the color, the title, and if needed the labels on the $x$- and $y$-axes.

```{r}
ggplot(data = nesarc.sub, aes(x = TotalCigsSmoked)) +
  geom_histogram(binwidth = 200, fill = "gray") + 
  labs(x = "Estimated Cigarettes Smoked per Month") +
  theme_bw()
```  

### Creating Density Plots

```{r}
ggplot(data = nesarc.sub, aes(x = TotalCigsSmoked)) +
  geom_density(fill = "sienna") + 
  labs(x = "Estimated Cigarettes Smoked per Month") +
  theme_bw()
```  

```{r}
ggplot(data = nesarc.sub, aes(x = DailyCigsSmoked)) +
  geom_density(fill = "slategrey") + 
  labs(x = "Reported Cigarettes Smoked per Day") +
  theme_bw()
```  

### The 3 S's

```{r}
summary(nesarc.sub$TotalCigsSmoked)
fivenum(nesarc.sub$TotalCigsSmoked)
median(nesarc.sub$TotalCigsSmoked, na.rm = TRUE)
IQR(nesarc.sub$TotalCigsSmoked, na.rm = TRUE)
```

The `shape` of the distribution for the estimated cigarettes smoked per month is skewed to the right.  The `center` (median) of the distribution is `r median(nesarc.sub$TotalCigsSmoked, na.rm = TRUE)` and the `spread` (interquartile range) for the distribution is `r IQR(nesarc.sub$TotalCigsSmoked, na.rm = TRUE)`.

### Bivariate Graphs (C $\rightarrow$ C)

Consider using depression status (`Depression`) to predict nicotine dependence (`TobaccoDependence`).

```{r}
T1 <- xtabs(~ TobaccoDependence + Depression, data = nesarc.sub)
T1
T2 <- prop.table(T1, 2)
T2
barplot(T2)
```

The last graph needs a legend as well as a title.  While it is possible to construct a legend and title for the last graph, it is much better to use an approach that will generate the legend automatically.

```{r}
g1 <- ggplot(data = nesarc.sub, aes(x = Depression, fill = TobaccoDependence)) +
  geom_bar(position = "fill") + 
  theme_bw()
g1
```

Tweaking the last graph.  

```{r}
g1 + labs(x = "", y = "", 
          title = "Fraction of young adult smokers with and without\n nicotine dependence by depression status") + 
  scale_fill_discrete(name="Tobacco Addiction\nStatus") 
```

```{r}
g2 <- ggplot(data = na.omit(nesarc.sub), aes(x = SmokingFreq, fill = TobaccoDependence)) +
  geom_bar(position = "fill") + 
  theme_bw() + 
  theme(axis.text.x  = element_text(angle = 75, vjust = 0.5)) +
  labs(x = "")
g2
 g3 <- g2 + facet_grid(Sex ~ .) +
  theme(axis.text.x  = element_text(angle = 85, vjust = 0.5)) +
  labs(x = "", y = "", title = "Fraction of young adult smokers with and without\n nicotine dependence by smoking frequency") + 
  scale_fill_discrete(name="Tobacco Addiction\nStatus")
 g3
```


```{r}
ggplot(data = na.omit(nesarc.sub), aes(x = CigsSmokedFac, fill = TobaccoDependence)) +
  geom_bar(position = "fill") + 
  theme_bw() + facet_grid(Sex ~ Depression) +
  theme(axis.text.x  = element_text(angle = 85, vjust = 0.5)) +
  labs(x = "Estimated Number of Cigarettes Smoked per Month", y = "", title = "Fraction of young adult smokers with and without\n nicotine dependence by smoking quantity") + 
  scale_fill_discrete(name="Tobacco Addiction\nStatus") 
```
 
### Using Mosaic Plots

```{r fig.width = 8, fig.height =8}
library(vcd)
mosaic(~TobaccoDependence + Depression, data = nesarc.sub) 
mosaic(~TobaccoDependence + Depression, data = nesarc.sub, shade = TRUE) 
mosaic(~TobaccoDependence + Depression + Sex, data = nesarc.sub, shade = TRUE)
mosaic(~TobaccoDependence + Depression + Sex + CigsSmokedFac, data = nesarc.sub, shade = TRUE)  
```

### Some Statistics

Note `nesarc.sub` is not the same data set discussed in the book.  In the book, the subset is smaller since it includes only people who have smoked every day (`S3AQ41`).  This subset has $(n = `r dim(nesarc.sub)[1]`)$ compared to the one in the book which has $n = 1320$. 

```{r}
ggplot(data = nesarc.sub, aes(x = DailyCigsSmoked)) + 
   geom_density(fill = "gray") + 
   facet_grid(Depression~.) +
   theme_bw()
tapply(nesarc.sub$DailyCigsSmoked, list(nesarc.sub$Depression), sd, na.rm = TRUE) 
 ggplot(data = nesarc.sub, aes(x = Depression, y = DailyCigsSmoked, fill = Depression)) + geom_violin() + 
   theme_bw()
```

```{r}
t.test(DailyCigsSmoked ~ Depression, data = nesarc.sub)
t.test(TotalCigsSmoked ~ Depression, data = nesarc.sub)
t.test(TotalCigsSmoked ~ Depression, data = nesarc.sub, var.equal = TRUE) # assumptions not really met here
summary(aov(TotalCigsSmoked ~ Depression, data = nesarc.sub))
# OR
mod <- lm(TotalCigsSmoked ~ Depression, data = nesarc.sub)
anova(mod)
``` 

### Taking a permuation approach

You should add your own notes here when I explain this in class.

```{r}
set.seed(123)
SIMS <- 1000 - 1
stat <- numeric(SIMS)
TS <- t.test(DailyCigsSmoked ~ Depression, data = nesarc.sub)$stat
for(i in 1:SIMS){
  stat[i] <- t.test(DailyCigsSmoked ~ sample(Depression), data = nesarc.sub)$stat
}
ggplot(data = data.frame(stat = stat), aes(x = stat)) + 
  geom_density(fill = "pink") + 
  theme_bw() + 
  stat_function(fun = dt, arg = list(df = 733), color = "red", linetype = "dashed") + geom_vline(x = c(TS, -TS), linetype = "dashed", color = "blue")
TS
pv <- (sum(stat >= abs(TS))*2 + 1) / SIMS
pv
# ANOVA
set.seed(123)
SIMS <- 1000 - 1
stat <- numeric(SIMS)
TSF <- anova(lm(DailyCigsSmoked ~ Depression, data = nesarc.sub))[1, 4]
for(i in 1:SIMS){
  stat[i] <- anova(lm(DailyCigsSmoked ~ sample(Depression), data = nesarc.sub))[1, 4]
}
ggplot(data = data.frame(stat = stat), aes(x = stat)) + 
  geom_density(fill = "pink") + 
  theme_bw() + 
  stat_function(fun = df, arg = list(df1 = 1, df2 = 1695), color = "red", linetype = "dashed") + geom_vline(x = TSF, linetype = "dashed", color = "blue") + ylim(0, 1)
TSF
pvF <- (sum(stat >= TSF) + 1) / SIMS
pvF
```

## Another Example

```{r}
ggplot(data = nesarc.sub, aes(x = Ethnicity, y = TotalCigsSmoked, fill = Ethnicity)) + 
  geom_boxplot() +
  guides(fill = FALSE) +
  theme_bw() + 
  labs(x = "", y = "Estimated Number of Cigarettes Smoked per Month\n")
ggplot(data = nesarc.sub, aes(x = Ethnicity, y = TotalCigsSmoked, fill = Ethnicity)) + 
  geom_violin() +
  guides(fill = FALSE) +
  theme_bw() + 
  labs(x = "", y = "Estimated Number of Cigarettes Smoked per Month\n")
```

$H_0:\mu_{\text{Caucassian}} = \mu_{\text{African American}}=\mu_{\text{Native American}} =\mu_{\text{Asian}} =\mu_{\text{Hispanic}}$ versus
$H_1:\mu_{i} \neq \mu_{j}$ for some $(i, j)$.

```{r}
eth.aov <- aov(TotalCigsSmoked ~ Ethnicity, data = nesarc.sub)
summary(eth.aov)
CIs <- TukeyHSD(eth.aov)
CIs
opar <- par(no.readonly = TRUE)  # copy of current settings
par(mar =c(1, 15, 2, 1)) 
plot(CIs, las = 1)
par(opar)                        # restore original settings
```

Evidence (*p* = `r summary(eth.aov)[[1]][1,5]`) suggests a strong relationship between ethnicity and the estimated number of cigarettes an individual smokes per month. Post hoc comparisons suggest mean differences in the number of cigarettes smoked per month for African American and Caucasian, Asian and Caucasian, and Hispanic and Caucasian.

## Chi Square Tests

```{r fig.width = 10, fig.height = 10}
library(vcd)
mosaic(~TobaccoDependence + Depression, data = nesarc.sub, shade = TRUE)
xtabs(~TobaccoDependence + Depression, data = nesarc.sub)
T1 <- prop.table(xtabs(~TobaccoDependence + Depression, data = nesarc.sub), 1)
T1
CR <- chisq.test(nesarc.sub$TobaccoDependence, nesarc.sub$Depression)
CR
```

A Chi Square test of independence suggest that among young adult smokers, those with past year nicotine dependence were more likely to have experiences major depression in their lifetime (`r T1[1, 2]*100`%) compared to those without past year nicotine dependence (`r T1[2, 2]*100`%), $\chi^2 = `r CR$stat`$, 1 df, *p* < 0.0001.

```{r}
mosaic(~TobaccoDependence + SmokingFreq, data = nesarc.sub, shade = TRUE)
TA <- xtabs(~TobaccoDependence + SmokingFreq, data = nesarc.sub)
TA
T1 <- prop.table(xtabs(~TobaccoDependence + SmokingFreq, data = nesarc.sub), 1)
T1
CR <- chisq.test(nesarc.sub$TobaccoDependence, nesarc.sub$SmokingFreq)
CR
```

The Chi Square test of independence suggest there is an association between smoking frequency and tobacco dependence.
To determine if there are significant differences among the six levels of smoking frequency requires one to examine $\binom{6}{2} =$ `r choose(6, 2)` possible differences with an appropriate $\alpha$ value to account for the multiple comparisons.  Use the Bonferroni approach where $\alpha = 0.05/(15) = `r 0.05/15` to control the experiment-wise error rate.

```{r}
chisq.test(TA[, c(1, 2)]) # not significant
chisq.test(TA[, c(1, 3)]) # not significant
chisq.test(TA[, c(1, 4)]) # significant
chisq.test(TA[, c(1, 5)]) # significant
chisq.test(TA[, c(1, 6)]) # significant
chisq.test(TA[, c(2, 3)]) # not significant
chisq.test(TA[, c(2, 4)]) # not significant
chisq.test(TA[, c(2, 5)]) # not significant 
chisq.test(TA[, c(2, 6)]) # significant
chisq.test(TA[, c(3, 4)]) # not significant
chisq.test(TA[, c(3, 5)]) # not significant
chisq.test(TA[, c(3, 6)]) # significant
chisq.test(TA[, c(4, 5)]) # not significant
chisq.test(TA[, c(4, 6)]) # significant
chisq.test(TA[, c(5, 6)]) # significant
```

There are significant differences between the percent of subjects that have nicotine dependence while smoking once a month or less and 3 to 4 Days/week, 5 to 6 Days/week, and Every Day.  There are significant differences between  percent of subjects that have nicotine dependence while smoking Every day and each of the less frequent smoking modes.


```{r results='asis', echo = TRUE}
library(moonBook)
out <- mytable(Depression + TobaccoDependence ~  Age + SmokingFreq + DailyCigsSmoked + Ethnicity + Sex +DaysSmoke + TotalCigsSmoked + CigsSmokedFac , data = nesarc.sub)
myhtml(out)
```